---
title: "Compute network plots of current connectivities"
---


```{r}
library(ggplot2)
library(MASS)
library(tidyverse) 
library(gdistance)
library(fields)
library(scales)
library(igraph)
library(ggraph)
library(tidygraph)
library(ggrepel)
library(reshape2)
library(GGally)
library(ggpubr)
library(DHARMa)
library(car)
library(glmmTMB)

source("Functions.R")
```


parameters for plotting

```{r}
my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))
```



Read directed matrices

```{r}
mat_median_min_singlepoints_gal <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_gal.rds"))
mat_median_min_singlepoints_can <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_can.rds"))
mat_median_min_singlepoints_az <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_az.rds"))

```


Read undirected matrices

```{r}

mat_medianMin_sym_Galapagos <- readRDS(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Galapagos.rds"))
mat_medianMin_sym_Canaries <- readRDS(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Canaries.rds"))
mat_medianMin_sym_Azores <- readRDS(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Azores.rds"))
```




## compute % littoral species in floras

```{r}
data_Galapagos <- read_csv(here::here("b_data/Data_Galapagos/data_galapagos.csv"), col_names = TRUE, trim_ws = TRUE)
data_Canaries <- read_csv(here::here("b_data/Data_Canaries/data_canaries.csv"), col_names = TRUE, trim_ws = TRUE)
data_Azores <- read_csv(here::here("b_data/Data_Azores/data_azores.csv"), col_names = TRUE)

df_Galapagos <- as.data.frame(data_Galapagos)
df_Galapagos <- unite(df_Galapagos, name_species, c(Genus, Species), remove=FALSE)
df_Galapagos$name_species <- gsub('Sesuvium_edmondstonei', 'Sesuvium_edmonstonei', df_Galapagos$name_species)
df_Galapagos$name_species <- gsub('Pycreus_polystachyos', 'Cyperus_polystachyos', df_Galapagos$name_species)

df_Canaries <- as.data.frame(data_Canaries)
df_Azores <- as.data.frame(data_Azores)

df_Azores <- df_Azores %>% 
  rename(
    Corvo = COR,
    Faial = FAI,
    Flores = FLO,
    Graciosa = GRA,
    Pico = PIC,
    SantaMaria = SMR,
    SaoJorge = SJG,
    SaoMiguel = SMG,
    Terceira = TER
  )
```

Filter lowland (azores database already filtered)

```{r}
df_Galapagos_lowland <- subset(df_Galapagos, df_Galapagos$habitats_lessthan500m %in% "1")
df_Canaries_lowland <- subset(df_Canaries, df_Canaries$habitats_lessthan500m %in% "1")
```


Galapagos

```{r}
# Define the columns for islands
island_cols <- c("Espanola", "Fernandina", "Floreana", "Genovesa", "Isabela", "Marchena", "Pinta", "Pinzon", "San_Cristobal", "Santa_Cruz", "Santa_Fe", "Santiago")

# Initialize a dataframe to store the results
results <- data.frame(
  Island = character(),
  Percent_Littoral = numeric(),
  Total_Species = integer(),
  stringsAsFactors = FALSE
)

# Iterate over each island
for (island in island_cols) {
  # Subset the species present in the island
  species_in_island <- df_Galapagos_lowland[complete.cases(df_Galapagos_lowland[, island]), ]
  
  # Calculate the total number of species in the island
  total_species <- nrow(species_in_island)
  
  # Calculate the percentage of littoral species
  littoral_species <- species_in_island[species_in_island$littoral %in% c("exclusive", "presence"), ]
  percent_littoral <- (nrow(littoral_species) / total_species) * 100
  
  # Add the results to the dataframe
  results <- rbind(results, data.frame(
    Island = island,
    Percent_Littoral = percent_littoral,
    Total_Species = total_species,
    stringsAsFactors = FALSE
  ))
}

df_percent_flora_gal <- results

```



Canaries

```{r}
# Define the columns for islands
island_cols <- c("El_Hierro", "La_Palma", "La_Gomera", "Tenerife", "Gran_Canaria", "Fuerteventura", "Lanzarote")

# Initialize a dataframe to store the results
results <- data.frame(
  Island = character(),
  Percent_Littoral = numeric(),
  Total_Species = integer(),
  stringsAsFactors = FALSE
)

# Iterate over each island
for (island in island_cols) {
  # Subset the species present in the island
  species_in_island <- df_Canaries_lowland[df_Canaries_lowland[[island]] == 1, ]
  
  # Calculate the total number of species in the island
  total_species <- nrow(species_in_island)
  
  # Calculate the percentage of littoral species
  littoral_species <- species_in_island[species_in_island$littoral %in% c("exclusive", "presence"), ]
  percent_littoral <- (nrow(littoral_species) / total_species) * 100
  
  # Add the results to the dataframe
  results <- rbind(results, data.frame(
    Island = island,
    Percent_Littoral = percent_littoral,
    Total_Species = total_species,
    stringsAsFactors = FALSE
  ))
}

df_percent_flora_can <- results


```


Azores

```{r}
# Define the columns for islands
island_cols <- c("SantaMaria", "SaoMiguel", "Terceira", "Graciosa", "SaoJorge", "Pico", "Faial", "Flores", "Corvo")

# Initialize a dataframe to store the results
results <- data.frame(
  Island = character(),
  Percent_Littoral = numeric(),
  Total_Species = integer(),
  stringsAsFactors = FALSE
)

# Iterate over each island
for (island in island_cols) {
  # Subset the species present in the island
  species_in_island <- df_Azores[df_Azores[[island]] == 1, ]
  
  # Calculate the total number of species in the island
  total_species <- nrow(species_in_island)
  
  # Calculate the percentage of littoral species
  littoral_species <- species_in_island[species_in_island$littoral %in% "1", ]
  percent_littoral <- (nrow(littoral_species) / total_species) * 100
  
  # Add the results to the dataframe
  results <- rbind(results, data.frame(
    Island = island,
    Percent_Littoral = percent_littoral,
    Total_Species = total_species,
    stringsAsFactors = FALSE
  ))
}

df_percent_flora_az <- results

```




Raw plots % littoral



```{r}



ymax_perc = max(c(df_percent_flora_gal$Percent_Littoral, df_percent_flora_can$Percent_Littoral, df_percent_flora_az$Percent_Littoral))


# Create plots with reordered dataframes
plot_gal <- ggplot(df_percent_flora_gal, aes(x = reorder(Island, -Percent_Littoral), y = Percent_Littoral)) +
  geom_bar(stat = "identity", fill = "gray25") +  # Create the bar plot
  geom_text(aes(label = Total_Species), 
            vjust = -0.5, 
            color = "black", 
            size = 4) +  # Add text labels for Total_Species at the top of each bar
  labs(x = NULL, 
       y = "% littoral species") +
  ylim(0, ymax_perc+0.8)+
  theme_classic() +  # Use a classic theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +# Rotate x-axis labels for better readability
  my_theme

plot_can <- ggplot(df_percent_flora_can, aes(x = reorder(Island, -Percent_Littoral), y = Percent_Littoral)) +
  geom_bar(stat = "identity", fill = "gray25") +  # Create the bar plot
  geom_text(aes(label = Total_Species), 
            vjust = -0.5, 
            color = "black", 
            size = 4) +  # Add text labels for Total_Species at the top of each bar
  labs(x = NULL, 
       y = "% littoral species") +
  ylim(0, ymax_perc+0.8)+
  theme_classic() +  # Use a classic theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better readability
my_theme

plot_az <- ggplot(df_percent_flora_az, aes(x = reorder(Island, -Percent_Littoral), y = Percent_Littoral)) +
  geom_bar(stat = "identity", fill = "gray25") +  # Create the bar plot
  geom_text(aes(label = Total_Species), 
            vjust = -0.5, 
            color = "black", 
            size = 4) +  # Add text labels for Total_Species at the top of each bar
  labs(x = NULL, 
       y = "% littoral species") +
  ylim(0, ymax_perc+0.8)+
  theme_classic() +  # Use a classic theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  my_theme


plot_percent_litt_islands <- ggarrange(
  plot_gal,
  plot_can,
  plot_az,
  
  labels = LETTERS[1:3],
  
  nrow = 3,
  ncol = 1
)

plot_percent_litt_islands

#ggsave(here::here("c_output/figures/plot_percent_litt_islands.png"), height = 10, width = 4.5)

```





## Plot networks


# % littoral ~ Degree

We use the inverted weights to consider the degree strenghts, as we are taking into account the sea currents network.

To compute betwennness, edges are interpreted as distances, not as connection strengths. Therefore, we need to use the true weights.

```{r}

compute_df_percent_flora_degree <- function(mat_min_conn_currents, df_percent) {
  
  g <- graph_from_adjacency_matrix(mat_min_conn_currents, mode = "directed", weighted = TRUE)
  
  # Create a data frame from the igraph object
  node_list <- data.frame(name = V(g)$name)
  
   # Merge the coordinates with the node list
  node_list <- merge(node_list, df_percent[,c("Island", "Percent_Littoral")], by.x = "name", by.y = "Island", all.x = TRUE)
  
  # Ensure the graph vertices match the island names
  V(g)$name <- as.character(V(g)$name)
  node_list$name <- as.character(node_list$name)
  
  # Convert the igraph object to a tidygraph object
  tg <- as_tbl_graph(g)
  
    # Add coordinates to the tidygraph object
  tg <- tg %>%
    activate(nodes) %>%
    left_join(node_list, by = c("name" = "name"))
  
    # Invert the weights: higher values become lower and vice versa
  E(tg)$inv_weight <- 1 / E(tg)$weight
  
  
  ### CENTRALITY MEASURES
  
  # in degree (sinks)
  V(tg)$in_degree <- strength(tg, mode = "in", weights = E(tg)$inv_weight)
  
  # out_degree (pulls)
  V(tg)$out_degree <- strength(tg, mode = "out", weights = E(tg)$inv_weight)
  
    # Closeness
  V(tg)$closeness <- closeness(tg, weights = E(tg)$weight)
  
  # Betwenness
  V(tg)$betweenness <- betweenness(tg, weights = E(tg)$weight)
  

  

  df_percent_litt_degree_az <- data.frame(tg)
  
  return(df_percent_litt_degree_az)
  
}

# Galapagos

df_percent_flora_gal[which(df_percent_flora_gal$Island == "San_Cristobal"), "Island"] <- "SanCristobal"
df_percent_flora_gal[which(df_percent_flora_gal$Island == "Santa_Cruz"), "Island"] <- "SantaCruz"
df_percent_flora_gal[which(df_percent_flora_gal$Island == "Santa_Fe"), "Island"] <- "SantaFe"

df_percent_degree_gal <- compute_df_percent_flora_degree(mat_median_min_singlepoints_gal, df_percent_flora_gal)
df_percent_degree_gal$archipelago <- rep("Galapagos", times = nrow(df_percent_degree_gal))


# Canaries
  
df_percent_flora_can[which(df_percent_flora_can$Island == "El_Hierro"), "Island"] <- "ElHierro"
df_percent_flora_can[which(df_percent_flora_can$Island == "Gran_Canaria"), "Island"] <- "GranCanaria"
df_percent_flora_can[which(df_percent_flora_can$Island == "La_Gomera"), "Island"] <- "LaGomera"
df_percent_flora_can[which(df_percent_flora_can$Island == "La_Palma"), "Island"] <- "LaPalma"


df_percent_degree_can <- compute_df_percent_flora_degree(mat_median_min_singlepoints_can, df_percent_flora_can)
df_percent_degree_can$archipelago <- rep("Canaries", times = nrow(df_percent_degree_can))


# Azores
  
df_percent_degree_az <-compute_df_percent_flora_degree(mat_median_min_singlepoints_az, df_percent_flora_az)
df_percent_degree_az$archipelago <- rep("Azores", times = nrow(df_percent_degree_az))



df_percent_degree_all <- rbind(df_percent_degree_gal, df_percent_degree_can, df_percent_degree_az)

# For readability:

df_percent_degree_all$closeness <- df_percent_degree_all$closeness*1000

df_percent_degree_all <- df_percent_degree_all %>% 
  rename(island = name)

options(scipen = 999)

df_percent_degree_all


#write.csv(df_percent_degree_all, here::here("c_output/tables/island_centrality_currents.csv"))


```



Stats centrality


```{r}

# Group the data by archipelago and summarize
summary_long <- df_percent_degree_all %>%
  group_by(archipelago) %>%
  summarise(
    Mean_Percent_Littoral = mean(Percent_Littoral),
    SD_Percent_Littoral = sd(Percent_Littoral),
    Mean_in_degree = mean(in_degree),
    SD_in_degree = sd(in_degree),
    Mean_out_degree = mean(out_degree),
    SD_out_degree = sd(out_degree),
        Mean_closeness = mean(closeness),
    SD_closeness = sd(closeness),
    Mean_betweenness = mean(betweenness),
    SD_betweenness = sd(betweenness),
  ) %>%
  pivot_longer(cols = -archipelago, names_to = "Metric", values_to = "Value")

# Spread the data to get archipelagos as columns
summary_wide <- summary_long %>%
  pivot_wider(names_from = archipelago, values_from = Value)

options(scipen = 999)

summary_wide

#write.csv(summary_wide, here::here("c_output/tables/stats_islands.csv"))


```







Plot correlations centrality measures


```{r}
df_percent_degree_all <- df_percent_degree_all %>% 
  rename(In_degree = in_degree,
         Out_degree = out_degree,
         Betweenness = betweenness,
         Closeness = closeness)


plot_cor_centrality_measures <- ggpairs(df_percent_degree_all, aes(color = archipelago),
        columns = 3:6,  # Specify the columns to plot
        lower = 'blank', diag = 'blank',
        upper = list(continuous = wrap("cor", size = 3))) + 
  scale_color_manual(
    values = c("#e31a1c", "#33a02c", "#1f78b4"))+
  theme_classic() +
  my_theme

plot_cor_centrality_measures

#ggsave(here::here("c_output/figures/plot_cor_centrality_measures.png"), height = 5, width = 6)

```





## Test effect of centrality on % littoral

```{r}
#cCompute proportion
df_percent_degree_all$Percent_Littoral_Prop <- df_percent_degree_all$Percent_Littoral / 100

```


### Degree

```{r}

# Log-transform the in_degree variable
df_percent_degree_all <- df_percent_degree_all %>%
  mutate(log_in_degree = log(In_degree))

# Center the log-transformed in_degree variable
df_percent_degree_all <- df_percent_degree_all %>%
  mutate(centered_log_in_degree = log_in_degree - mean(log_in_degree, na.rm = TRUE))

# Model

model_degree <- glmmTMB(Percent_Littoral_Prop ~ archipelago*centered_log_in_degree, data = df_percent_degree_all, family = beta_family(link = "logit"))

# Result

summary(model_degree)

car::Anova(model_degree)

```

Model diagnostics

```{r}
simulationOutput_surv_d<-simulateResiduals(model_degree) 
dharma_plot_surv<-plot(simulationOutput_surv_d)
surv_dharma<-testResiduals(simulationOutput_surv_d)
```


Save data

```{r}
# Extract summary information
summary_info <- summary(model_degree)

# Extract coefficients and their statistics
coefficients_df <- summary_info$coefficients

# Save coefficients to CSV
write.csv(coefficients_df$cond, file = here::here("c_output/tables/model_degree_summary.csv"), row.names = TRUE)

write.csv(as.data.frame(car::Anova(model_degree)), file = here::here("c_output/tables/model_degree_anova.csv"), row.names = TRUE)
```

### Closeness

```{r}

# come back to original values

df_percent_degree_all$closeness <- df_percent_degree_all$closeness/1000

# scale them

df_percent_degree_all$Closeness_scaled <- scale(df_percent_degree_all$Closeness)


# Model

model_closeness <- glmmTMB(Percent_Littoral_Prop ~ archipelago*Closeness_scaled, data = df_percent_degree_all, family = beta_family(link = "logit"))

# Result

summary(model_closeness)

car::Anova(model_closeness)


```



```{r}
simulationOutput_surv_d<-simulateResiduals(model_closeness) 
dharma_plot_surv<-plot(simulationOutput_surv_d)
surv_dharma<-testResiduals(simulationOutput_surv_d)
```

Save data

```{r}
# Extract summary information
summary_info <- summary(model_closeness)

# Extract coefficients and their statistics
coefficients_df <- summary_info$coefficients

# Save coefficients to CSV
write.csv(coefficients_df$cond, file = here::here("c_output/tables/model_closeness_summary.csv"), row.names = TRUE)

write.csv(as.data.frame(car::Anova(model_closeness)), file = here::here("c_output/tables/model_closeness_anova.csv"), row.names = TRUE)
```




### Betwenness

```{r}
# Model

df_percent_degree_all$Betweenness[which(df_percent_degree_all$Betweenness == 0)] <- 0.01

model_betwenness <- glmmTMB(Percent_Littoral_Prop ~ archipelago*Betweenness, data = df_percent_degree_all, family = beta_family(link = "logit"))

# Result

summary(model_betwenness)

car::Anova(model_betwenness)


```



```{r}
simulationOutput_surv_d<-simulateResiduals(model_betwenness) 
dharma_plot_surv<-plot(simulationOutput_surv_d)
surv_dharma<-testResiduals(simulationOutput_surv_d)
```

Save data

```{r}
# Extract summary information
summary_info <- summary(model_betwenness)

# Extract coefficients and their statistics
coefficients_df <- summary_info$coefficients

# Save coefficients to CSV
write.csv(coefficients_df$cond, file = here::here("c_output/tables/model_betwenness_summary.csv"), row.names = TRUE)

write.csv(as.data.frame(car::Anova(model_betwenness)), file = here::here("c_output/tables/model_betwenness_anova.csv"), row.names = TRUE)
```




















## Plot correlations % littoral and centrality measures


```{r}
coord_centroids_gal <- read.csv(here::here("b_data/Data_Galapagos/centroids_network_plot.csv"))
coord_centroids_can <- read.csv(here::here("b_data/Data_Canaries/centroids_network_plot.csv"))
coord_centroids_az <- read.csv(here::here("b_data/Data_Azores/centroids_network_plot.csv"))
```



```{r}
back_transform_log <- function(x) {
  return(exp(x))  # Back transform using exponential function (inverse of log)
}

# scale closeness for axis readability:

df_percent_degree_all$closeness <- df_percent_degree_all$closeness/1000

```


```{r}
# Define the mathematical symbols for labels
in_degree_label <- expression(C[In]^W)
closeness_label <- expression(C[C])
betweenness_label <- expression(C[B])
```



% Littoral & centrality

```{r}

# maximum %

max_perc_litt <- max(df_percent_degree_all$Percent_Littoral)



# Galapagos 

plot_litt_degree_gal <- ggplot(df_percent_degree_all[df_percent_degree_all$archipelago == "Galapagos",], aes(x = in_degree, y = Percent_Littoral)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
  labs(x = in_degree_label, y = "% littoral plants in lowland") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


plot_litt_cc_gal <- ggplot(df_percent_degree_all[df_percent_degree_all$archipelago == "Galapagos",], aes(x = closeness, y = Percent_Littoral)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
      scale_x_continuous(labels = function(x) sprintf("%.2f", back_transform_log(x))) +
  labs(x = closeness_label, y = "% littoral plants in lowland") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


# Canaries 

plot_litt_degree_can <- ggplot(df_percent_degree_all[df_percent_degree_all$archipelago == "Canaries",], aes(x = log(in_degree), y = Percent_Littoral)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "red") + # Add trend lines
   scale_x_continuous(labels = function(x) sprintf("%.2f", back_transform_log(x))) +
  labs(x = in_degree_label, y = "% littoral plants in lowland") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


plot_litt_cc_can <- ggplot(df_percent_degree_all[df_percent_degree_all$archipelago == "Canaries",], aes(x = closeness, y = Percent_Littoral)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "red") + # Add trend lines
      scale_x_continuous(labels = function(x) sprintf("%.3f", back_transform_log(x))) +
  labs(x = closeness_label, y = "% littoral plants in lowland") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


# Azores 

plot_litt_degree_az <- ggplot(df_percent_degree_all[df_percent_degree_all$archipelago == "Azores",], aes(x = in_degree, y = Percent_Littoral)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
  labs(x = in_degree_label, y = "% littoral plants in lowland") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


plot_litt_cc_az <- ggplot(df_percent_degree_all[df_percent_degree_all$archipelago == "Azores",], aes(x = closeness, y = Percent_Littoral)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
      scale_x_continuous(labels = function(x) sprintf("%.2f", back_transform_log(x))) +
  labs(x = closeness_label, y = "% littoral plants in lowland") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme




plot_perc_litt_centrality <- ggarrange(
  plot_litt_degree_gal+ labs(title = ""), plot_litt_degree_can + ylab(NULL)+ labs(title = ""), plot_litt_degree_az + ylab(NULL)+ labs(title = ""),
  plot_litt_cc_gal+ labs(title = ""), plot_litt_cc_can + ylab(NULL)+ labs(title = ""), plot_litt_cc_az + ylab(NULL)+ labs(title = ""),
  
  ncol = 3,
  nrow = 2,
  
  labels = LETTERS[1:6]
)


plot_perc_litt_centrality

ggsave(here::here("c_output/figures/plot_perc_litt_centrality.png"), height = 9, width = 9)

```




# NETWORKS


function to plot the % litt in undirected networks




## Undirected networks and % littoral flora


Currents

```{r}



# Galapagos

df_percent_flora_gal <- df_percent_flora_gal %>% 
  rename(island = Island)

plot_network_litt_gal <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Galapagos,
                      centroids = coord_centroids_gal,
                      node_size = 8,
                      percent_littoral = df_percent_flora_gal[,c(1,2)],
                      scale = "none")

plot_network_litt_gal

ggsave(here::here("c_output/figures/network_littoral_gal.png"), height = 8, width = 9)


# Canaries

df_percent_flora_can <- df_percent_flora_can %>% 
  rename(island = Island)

plot_network_litt_can <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Canaries,
                      centroids = coord_centroids_can,
                      node_size = 8,
                      percent_littoral = df_percent_flora_can[,c(1,2)],
                      scale = "log_scale")

plot_network_litt_can

ggsave(here::here("c_output/figures/network_littoral_can.png"), height = 8, width = 9)


# Azores

df_percent_flora_az <- df_percent_flora_az %>% 
  rename(island = Island)

plot_network_litt_az <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Azores,
                      centroids = coord_centroids_az,
                      node_size = 8,
                      percent_littoral = df_percent_flora_az[,c(1,2)],
                      scale = "sqrt_scale")

plot_network_litt_az

#ggsave(here::here("c_output/figures/network_littoral_az.png"), height = 8, width = 9)

```


Plot combined network littoral

```{r}


plot_network_litt_all <- ggarrange(
  plot_network_litt_gal,
  plot_network_litt_can + xlab(""),
  plot_network_litt_az+ xlab(""),
  
  labels = LETTERS[1:3],
  
  ncol = 1,
  nrow = 3
)

plot_network_litt_all

#ggsave(here::here("c_output/figures/plot_network_litt_all.png"), height = 12, width = 8)



```


Geo, age, area

```{r}
# Galapagos

df_percent_flora_gal <- df_percent_flora_gal %>% 
  rename(island = Island)

plot_network_geo_gal <- plot_network_littoral(weighted_matrix = mat_geo_gal,
                      centroids = coord_centroids_gal,
                      node_size = 8,
                      percent_littoral = df_percent_flora_gal[,c(1,2)],
                      scale = "none")

plot_network_age_gal <- plot_network_littoral(weighted_matrix = mat_age_gal,
                      centroids = coord_centroids_gal,
                      node_size = 8,
                      percent_littoral = df_percent_flora_gal[,c(1,2)],
                      scale = "none")

plot_network_area_gal <- plot_network_littoral(weighted_matrix = mat_area_gal,
                      centroids = coord_centroids_gal,
                      node_size = 8,
                      percent_littoral = df_percent_flora_gal[,c(1,2)],
                      scale = "log")

# Canaries

df_percent_flora_can <- df_percent_flora_can %>% 
  rename(island = Island)

plot_network_litt_can <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Canaries,
                      centroids = coord_centroids_can,
                      node_size = 8,
                      percent_littoral = df_percent_flora_can[,c(1,2)],
                      scale = "log_scale")


# Azores

df_percent_flora_az <- df_percent_flora_az %>% 
  rename(island = Island)

plot_network_litt_az <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Azores,
                      centroids = coord_centroids_az,
                      node_size = 8,
                      percent_littoral = df_percent_flora_az[,c(1,2)],
                      scale = "sqrt_scale")

```





## Directed networks & centrality


Get in-degrees

```{r}

df_degree_gal <- df_percent_degree_gal[,c("name","in_degree")] %>% 
  rename(
    island = name
  )

df_degree_can <- df_percent_degree_can[,c("name","in_degree")] %>% 
  rename(
    island = name
  )

df_degree_az <- df_percent_degree_az[,c("name","in_degree")] %>% 
  rename(
    island = name
  )
```


Plots

```{r}

# Galapagos

df_percent_flora_gal <- df_percent_flora_gal %>% 
  rename(island = Island)

plot_network_degree_gal <- plot_network_in_degree(weighted_matrix = mat_medianMin_sym_Galapagos, 
                       centroids = coord_centroids_gal, 
                       df_degree = df_degree_gal,
                       node_size = 8,
                      percent_littoral = df_percent_flora_gal[,c(1,3)],
                      scale_edge = "none",
                      scale_degree = "none",
                      legend_label = in_degree_label)

plot_network_degree_gal

#ggsave(here::here("c_output/figures/network_degree_gal.png"), height = 8, width = 9)



# Canaries

df_percent_flora_can <- df_percent_flora_can %>% 
  rename(island = Island)


plot_network_degree_can <- plot_network_in_degree(weighted_matrix = mat_medianMin_sym_Canaries, 
                       centroids = coord_centroids_can, 
                       df_degree = df_degree_can,
                       node_size = 8,
                      percent_littoral = df_percent_flora_can[,c(1,3)],
                      scale_edge = "log_scale",
                      scale_degree = "log_scale",
                      legend_label = in_degree_label)+
    scale_color_viridis(
    name = "In-degree centrality", 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))
  

plot_network_degree_can


#ggsave(here::here("c_output/figures/network_degree_can.png"), height = 8, width = 9)


# Azores

df_percent_flora_az <- df_percent_flora_az %>% 
  rename(island = Island)


plot_network_degree_az <- plot_network_in_degree(weighted_matrix = mat_medianMin_sym_Azores, 
                       centroids = coord_centroids_az, 
                       df_degree = df_degree_az,
                       node_size = 8,
                      percent_littoral = df_percent_flora_az[,c(1,3)],
                      scale_edge = "sqrt_scale",
                      scale_degree = "none",
                      legend_label = in_degree_label)

plot_network_degree_az


#ggsave(here::here("c_output/figures/network_degree_az.png"), height = 8, width = 9)


```




Plot combined network & regression degree

```{r}


plot_regr_centrality_gal <- ggarrange(
  
  plot_litt_degree_gal + ylab(NULL), NULL,
plot_litt_cc_gal+ ylab(NULL), NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_gal <- annotate_figure(
  plot_regr_centrality_gal,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


plot_regr_centrality_can <- ggarrange(
  
  plot_litt_degree_can+ ylab(NULL), NULL,
plot_litt_cc_can+ ylab(NULL), NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_can <- annotate_figure(
  plot_regr_centrality_can,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


plot_regr_centrality_az <- ggarrange(
  
  plot_litt_degree_az+ ylab(NULL), NULL,
plot_litt_cc_az+ ylab(NULL),NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_az <- annotate_figure(
  plot_regr_centrality_az,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


#############################




plot_network_regr_degree <- ggarrange(
  
      plot_network_degree_gal + theme(legend.position = "left") +
    scale_color_viridis(name = in_degree_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) 
  , plot_regr_centrality_gal,
  
    plot_network_degree_can + theme(legend.position = "left") +
    scale_color_viridis(
    name = in_degree_label, 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) 
  , plot_regr_centrality_can,
  
  plot_network_degree_az + theme(legend.position = "left") +
    scale_color_viridis(name =in_degree_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) 
  , plot_regr_centrality_az,
  
  ncol = 2,
  nrow = 3,
  
  labels = LETTERS[1:6],
  
  widths = c(2, 1)
)

plot_network_regr_degree

ggsave(here::here("c_output/figures/plot_network_regr_degree.png"), height = 12, width = 12)


```








