---
title: "Mantel test Galapagos - Floristic & Currents conectivity"
---

In this script, we will convert the obtained floristic and currents matrices (floristic dissimilatiry matrix and oceanic current connectivity matrix) into distance matrices and perform a Procrustes analyses to explore their correlation and answer whether highly sea-current-connected islands more floristicaly similar. 

We will also analyze whether floristic similarities are related to geographic, area, and age distance between islands.

```{r}
library(ade4)
library(vegan)
library(tidyverse)
library(cluster)
library(dplyr)
library(tidyverse)
library(cluster)
library(ggplot2)
library(reshape2)
library(viridis)
library(ggpubr)
source(here::here("a_script/Functions.R"))
```

Parameters for plotting

```{r}
my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))
```


#### Load floristic distances


```{r}

# Galapagos

dist_mat_all_litt_gal <- as.matrix(read_rds(here::here("c_output/floristic_dist_matrices/dist_mat_all_litt_gal.rds")))
dist_mat_all_nolitt_gal <- as.matrix(read_rds(here::here("c_output/floristic_dist_matrices/dist_mat_all_nolitt_gal.rds")))

# Canaries

dist_mat_all_litt_can <- as.matrix(read_rds(here::here("c_output/floristic_dist_matrices/dist_mat_all_litt_can.rds")))
dist_mat_all_nolitt_can <- as.matrix(read_rds(here::here("c_output/floristic_dist_matrices/dist_mat_all_nolitt_can.rds")))

# Azores

dist_mat_all_litt_az <- as.matrix(read_rds(here::here("c_output/floristic_dist_matrices/dist_mat_all_litt_az.rds")))
dist_mat_all_nolitt_az <- as.matrix(read_rds(here::here("c_output/floristic_dist_matrices/dist_mat_all_nolitt_az.rds")))

```



Order rows and columns and convert matrices to distance objects

```{r}

# Galapagos


dist_mat_all_litt_gal <- dist_mat_all_litt_gal[order(rownames(dist_mat_all_litt_gal)), order(colnames(dist_mat_all_litt_gal))]
dist_mat_all_nolitt_gal <- dist_mat_all_nolitt_gal[order(rownames(dist_mat_all_nolitt_gal)), order(colnames(dist_mat_all_nolitt_gal))]

# Canaries


dist_mat_all_litt_can <- dist_mat_all_litt_can[order(rownames(dist_mat_all_litt_can)), order(colnames(dist_mat_all_litt_can))]
dist_mat_all_nolitt_can <- dist_mat_all_nolitt_can[order(rownames(dist_mat_all_nolitt_can)), order(colnames(dist_mat_all_nolitt_can))]


# Azores


dist_mat_all_litt_az <- dist_mat_all_litt_az[order(rownames(dist_mat_all_litt_az)), order(colnames(dist_mat_all_litt_az))]
dist_mat_all_nolitt_az <- dist_mat_all_nolitt_az[order(rownames(dist_mat_all_nolitt_az)), order(colnames(dist_mat_all_nolitt_az))]
```


#### Load current distances

```{r}
mat_medianMin_sym_Galapagos <- read_rds(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Galapagos.rds"))
mat_medianMin_sym_Canaries <- read_rds(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Canaries.rds"))
mat_medianMin_sym_Azores <- read_rds(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Azores.rds"))
```




#### Load currents geographic matrices and convert to distance objects [TODO]

```{r}

mat_geo_gal <- read.csv(here::here("b_data/Data_Galapagos/geographic_distance_galap.csv"))
rownames(mat_geo_gal) <- mat_geo_gal[,1]
mat_geo_gal<-mat_geo_gal[,-1]

mat_geo_can <- read.csv(here::here("b_data/Data_Canaries/geographic_distance_canaries.csv"))
rownames(mat_geo_can) <- mat_geo_can[,1]
mat_geo_can<-mat_geo_can[,-1]

mat_geo_az <- read.csv(here::here("b_data/Data_Azores/geographic_distance_azores.csv"))
rownames(mat_geo_az) <- mat_geo_az[,1]
mat_geo_az<-mat_geo_az[,-1]


```


Order names and convert to matrices

```{r}
mat_geo_gal <- as.matrix(mat_geo_gal[order(rownames(mat_geo_gal)), order(colnames(mat_geo_gal))])

mat_geo_can <- as.matrix(mat_geo_can[order(rownames(mat_geo_can)), order(colnames(mat_geo_can))])

mat_geo_az <- as.matrix(mat_geo_az[order(rownames(mat_geo_az)), order(colnames(mat_geo_az))])

```




#### Load island area matrices and convert to distance objects

Area

```{r}
df_area<-read.csv(here::here("b_data/Data_Galapagos/area_islands_galap.csv"))
df_area1<-column_to_rownames(df_area, 'X')

mat_area_gal <-as.matrix(daisy(df_area1, metric="euclidean"))



df_area<-read.csv(here::here("b_data/Data_Canaries/area_islands_canaries.csv"))
df_area1<-column_to_rownames(df_area, 'X')

mat_area_can <-as.matrix(daisy(df_area1, metric="euclidean"))


df_area<-read.csv(here::here("b_data/Data_Azores/area_islands_azores.csv"))
df_area1<-column_to_rownames(df_area, 'X')

mat_area_az <-as.matrix(daisy(df_area1, metric="euclidean"))
```



Ages

```{r}
df_age<-read.csv(here::here("b_data/Data_Galapagos/ages_islands_galap.csv"), sep = ";")
df_age1<-column_to_rownames(df_age, 'island')

mat_age_gal <-as.matrix(daisy(df_age1, metric="euclidean"))



df_age<-read.csv(here::here("b_data/Data_Canaries/ages_islands_canaries.csv"), sep = ";")
df_age1<-column_to_rownames(df_age, 'island')

mat_age_can <-as.matrix(daisy(df_age1, metric="euclidean"))


df_age<-read.csv(here::here("b_data/Data_Azores/ages_islands_azores.csv"), sep = ";")
df_age1<-column_to_rownames(df_age, 'island')

mat_age_az <-as.matrix(daisy(df_age1, metric="euclidean"))
```


# Procrustes test

```{r}

# Galapagos


procrustes_list_all_litt_gal <- perform_procrustes_analysis(mat_medianMin_sym_Galapagos, dist_mat_all_litt_gal)
procrustes_list_all_nolitt_gal <- perform_procrustes_analysis(mat_medianMin_sym_Galapagos, dist_mat_all_nolitt_gal)

procrustes_geo_list_all_litt_gal <- perform_procrustes_analysis(mat_geo_gal, dist_mat_all_litt_gal)
procrustes_geo_list_all_nolitt_gal <- perform_procrustes_analysis(mat_geo_gal, dist_mat_all_nolitt_gal)

procrustes_age_list_all_litt_gal <- perform_procrustes_analysis(mat_age_gal, dist_mat_all_litt_gal)
procrustes_age_list_all_nolitt_gal <- perform_procrustes_analysis(mat_age_gal, dist_mat_all_nolitt_gal)

procrustes_area_list_all_litt_gal <- perform_procrustes_analysis(mat_area_gal, dist_mat_all_litt_gal)
procrustes_area_list_all_nolitt_gal <- perform_procrustes_analysis(mat_area_gal, dist_mat_all_nolitt_gal)


# Canaries

procrustes_list_all_litt_can <- perform_procrustes_analysis(mat_medianMin_sym_Canaries, dist_mat_all_litt_can)
procrustes_list_all_nolitt_can <- perform_procrustes_analysis(mat_medianMin_sym_Canaries, dist_mat_all_nolitt_can)

procrustes_geo_list_all_litt_can <- perform_procrustes_analysis(mat_geo_can, dist_mat_all_litt_can)
procrustes_geo_list_all_nolitt_can <- perform_procrustes_analysis(mat_geo_can, dist_mat_all_nolitt_can)

procrustes_age_list_all_litt_can <- perform_procrustes_analysis(mat_age_can, dist_mat_all_litt_can)
procrustes_age_list_all_nolitt_can <- perform_procrustes_analysis(mat_age_can, dist_mat_all_nolitt_can)

procrustes_area_list_all_litt_can <- perform_procrustes_analysis(mat_area_can, dist_mat_all_litt_can)
procrustes_area_list_all_nolitt_can <- perform_procrustes_analysis(mat_area_can, dist_mat_all_nolitt_can)



# Azores

procrustes_list_all_litt_az <- perform_procrustes_analysis(mat_medianMin_sym_Azores, dist_mat_all_litt_az)
procrustes_list_all_nolitt_az <- perform_procrustes_analysis(mat_medianMin_sym_Azores, dist_mat_all_nolitt_az)

procrustes_geo_list_all_litt_az <- perform_procrustes_analysis(mat_geo_az, dist_mat_all_litt_az)
procrustes_geo_list_all_nolitt_az <- perform_procrustes_analysis(mat_geo_az, dist_mat_all_nolitt_az)

procrustes_age_list_all_litt_az <- perform_procrustes_analysis(mat_age_az, dist_mat_all_litt_az)
procrustes_age_list_all_nolitt_az <- perform_procrustes_analysis(mat_age_az, dist_mat_all_nolitt_az)

procrustes_area_list_all_litt_az <- perform_procrustes_analysis(mat_area_az, dist_mat_all_litt_az)
procrustes_area_list_all_nolitt_az <- perform_procrustes_analysis(mat_area_az, dist_mat_all_nolitt_az)

```



Currents - geo - age - are

```{r}
# Galapagos

procrustes_curr_geo_gal <- perform_procrustes_analysis(mat_geo_gal, mat_medianMin_sym_Galapagos)
procrustes_curr_age_gal <- perform_procrustes_analysis(mat_age_gal, mat_medianMin_sym_Galapagos)
procrustes_curr_area_gal <- perform_procrustes_analysis(mat_area_gal, mat_medianMin_sym_Galapagos)

# Canaries

procrustes_curr_geo_can <- perform_procrustes_analysis(mat_geo_can, mat_medianMin_sym_Canaries)
procrustes_curr_age_can <- perform_procrustes_analysis(mat_age_can, mat_medianMin_sym_Canaries)
procrustes_curr_area_can <- perform_procrustes_analysis(mat_area_can, mat_medianMin_sym_Canaries)


# Azores

procrustes_curr_geo_az <- perform_procrustes_analysis(mat_geo_az, mat_medianMin_sym_Azores)
procrustes_curr_age_az <- perform_procrustes_analysis(mat_age_az, mat_medianMin_sym_Azores)
procrustes_curr_area_az <- perform_procrustes_analysis(mat_area_az, mat_medianMin_sym_Azores)


df_cor_currents_factors_gal <- data.frame(
  
  "Island" = "Galapagos",
  
  "Geographic space" = procrustes_curr_geo_gal$procrustes_corr$t0,
  
  "Island age" = procrustes_curr_age_gal$procrustes_corr$t0,
  
  "Island area" = procrustes_curr_area_gal$procrustes_corr$t0
)

df_cor_currents_factors_can <- data.frame(
  
  "Island" = "Canaries",
  
  "Geographic space" = procrustes_curr_geo_can$procrustes_corr$t0,
  
  "Island age" = procrustes_curr_age_can$procrustes_corr$t0,
  
  "Island area" = procrustes_curr_area_can$procrustes_corr$t0
)

df_cor_currents_factors_az <- data.frame(
  
  "Island" = "Azores",
  
  "Geographic space" = procrustes_curr_geo_az$procrustes_corr$t0,
  
  "Island age" = procrustes_curr_age_az$procrustes_corr$t0,
  
  "Island area" = procrustes_curr_area_az$procrustes_corr$t0
)

df_cor_curr_factors <- rbind(df_cor_currents_factors_gal, df_cor_currents_factors_can, df_cor_currents_factors_az)

```

Plot results

```{r}
# Convert Island and Flora columns to factors with specified levels
df_cor_curr_factors$Island <- factor(df_cor_curr_factors$Island, levels = c("Galapagos", "Canaries", "Azores"))

# Reshape data to long format
df_long <- melt(df_cor_curr_factors, id.vars = "Island", variable.name = "Variable", value.name = "Correlation")



# Create the plot
plot_cor_currents_factors <- ggplot(df_long, aes(x = Variable, y = Island)) +
  geom_tile(aes(fill = Correlation), color = "white") +
  geom_text(aes(label = round(Correlation, 2)), color = "white", size = 4) +
  scale_fill_viridis_c(option = "turbo", limits = c(0, 1)) +  # Set limits for the color scale
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10)) +
  labs(fill = "Correlation",
       y = "Currents connectivity",
       x = "Factor connectivity")

plot_cor_currents_factors

ggsave(here::here("c_output/figures/corr_currents_factors_heatmap.png"), height = 3, width = 6)
```



```{r}


df_cor_gal <- data.frame(
  
  "Island" = rep("Galapagos", 2),
  
  "Flora" = c("Littoral", "Non-littoral"),
  
  "Currents" = c(procrustes_list_all_litt_gal$procrustes_corr$t0,
                 procrustes_list_all_nolitt_gal$procrustes_corr$t0),
  
  "Geographic space" = c(procrustes_geo_list_all_litt_gal$procrustes_corr$t0,
                         procrustes_geo_list_all_nolitt_gal$procrustes_corr$t0),
  
  "Island age" = c(procrustes_age_list_all_litt_gal$procrustes_corr$t0,
                   procrustes_age_list_all_nolitt_gal$procrustes_corr$t0),
  
  "Island area" = c(procrustes_area_list_all_litt_gal$procrustes_corr$t0,
                    procrustes_area_list_all_nolitt_gal$procrustes_corr$t0)
)

df_cor_can <- data.frame(
  
  "Island" = rep("Canaries", 2),
  
  "Flora" = c("Littoral", "Non-littoral"),
  
  "Currents" = c(procrustes_list_all_litt_can$procrustes_corr$t0,
                 procrustes_list_all_nolitt_can$procrustes_corr$t0),
  
  "Geographic space" = c(procrustes_geo_list_all_litt_can$procrustes_corr$t0,
                         procrustes_geo_list_all_nolitt_can$procrustes_corr$t0),
  
  "Island age" = c(procrustes_age_list_all_litt_can$procrustes_corr$t0,
                   procrustes_age_list_all_nolitt_can$procrustes_corr$t0),
  
  "Island area" = c(procrustes_area_list_all_litt_can$procrustes_corr$t0,
                    procrustes_area_list_all_nolitt_can$procrustes_corr$t0)
)

df_cor_az <- data.frame(
  
  "Island" = rep("Azores", 2),
  
  "Flora" = c("Littoral", "Non-littoral"),
  
  "Currents" = c(procrustes_list_all_litt_az$procrustes_corr$t0,
                 procrustes_list_all_nolitt_az$procrustes_corr$t0),
  
  "Geographic space" = c(procrustes_geo_list_all_litt_az$procrustes_corr$t0,
                         procrustes_geo_list_all_nolitt_az$procrustes_corr$t0),
  
  "Island age" = c(procrustes_age_list_all_litt_az$procrustes_corr$t0,
                   procrustes_age_list_all_nolitt_az$procrustes_corr$t0),
  
  "Island area" = c(procrustes_area_list_all_litt_az$procrustes_corr$t0,
                    procrustes_area_list_all_nolitt_az$procrustes_corr$t0)
)


df_cor <- rbind(df_cor_gal, df_cor_can, df_cor_az)


write.csv(df_cor, here::here("c_output/tables/correlation_table_results.csv"))

```



Plot results



```{r}

# Convert Island and Flora columns to factors with specified levels
df_cor$Island <- factor(df_cor$Island, levels = c("Galapagos", "Canaries", "Azores"))
df_cor$Flora <- factor(df_cor$Flora, levels = c("Non-littoral", "Littoral"))

# Reshape data to long format
df_long <- melt(df_cor, id.vars = c("Island", "Flora"), variable.name = "Variable", value.name = "Correlation")



# Create the plot
plot_cor_floras <- ggplot(df_long, aes(x = Variable, y = Flora)) +
  geom_tile(aes(fill = Correlation), color = "white") +
  geom_text(aes(label = round(Correlation, 2)), color = "white", size = 4) +
  scale_fill_viridis_c(option = "turbo", limits = c(0, 1)) +
  facet_wrap(~ Island) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10)) +
  labs(fill = "Correlation",
       y = "Floristic connectivity",
       x = "Factor connectivity")

plot_cor_floras

ggsave(here::here("c_output/figures/corr_floras_heatmap.png"), height = 4, width = 8)
```



```{r}


plot_cor_currents_factors1 <- ggarrange(
  NULL,
  plot_cor_currents_factors,
  NULL,
  nrow = 1,
  ncol = 3,
  widths = c(0.9, 1.8, 0.9)
)

plot_cor_currents_factors1 <- annotate_figure(
  plot_cor_currents_factors1,
  top = text_grob("", size = 16)
)

ggarrange(
  plot_cor_floras,
plot_cor_currents_factors1,

nrow = 2,
ncol = 1,

labels = LETTERS[1:2]

)


ggsave(here::here("c_output/figures/correlations_all.png"), height = 8, width = 8)

```




## sensitivity analysis

```{r}
df_Galapagos_sensitivity <- readRDS(here::here("c_output/sensitivity/df_Galapagos_sensitivity.rds"))
df_Canaries_sensitivity <- readRDS(here::here("c_output/sensitivity/df_Canaries_sensitivity.rds"))
df_Azores_sensitivity <- readRDS(here::here("c_output/sensitivity/df_Azores_sensitivity.rds"))


plot_sensitivity_gal <- perform_sensitivity_analysis(df_Galapagos_sensitivity, mat_medianMin_sym_Galapagos, min = 15, max = nrow(df_Galapagos_sensitivity))

plot_sensitivity_can <- perform_sensitivity_analysis(df_Canaries_sensitivity, mat_medianMin_sym_Canaries, min = 30, max = nrow(df_Canaries_sensitivity))

plot_sensitivity_az <- perform_sensitivity_analysis(df_Azores_sensitivity, mat_medianMin_sym_Azores, min = 15, max = nrow(df_Azores_sensitivity))


ggarrange(
  plot_sensitivity_gal,
  plot_sensitivity_can,
  plot_sensitivity_az,
  
  ncol = 3,
  nrow = 1,
  
  labels = LETTERS[1:3]
)

```


## Plot Procrustes

```{r}



# Kind 1


  png(here::here("c_output/figures/procrustes_kind1.png"), width = 25, height = 26, units = "cm", res = 600) 
  par(mfrow = c(3, 2))
  
  
  # Plot A: Procrustes plot for Littoral
  plot(procrustes_list_all_litt_gal$procrustes_result, kind = 1, type = "text", cex = 1, main = "Littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "")
  
  # Plot B: Procrustes plot for Non-littoral
  plot(procrustes_list_all_nolitt_gal$procrustes_result, kind = 1, type = "text", cex = 1, main = "Non-littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "", ylab = "")
  
    # Plot C: Procrustes plot for Littoral
  plot(procrustes_list_all_litt_can$procrustes_result, kind = 1, type = "text", cex = 1, main = "Littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "")
  
  # Plot D: Procrustes plot for Non-littoral
  plot(procrustes_list_all_nolitt_can$procrustes_result, kind = 1, type = "text", cex = 1, main = "Non-littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "", ylab = "")
  
    # Plot A: Procrustes plot for Littoral
  plot(procrustes_list_all_litt_az$procrustes_result, kind = 1, type = "text", cex = 1, main = "Littoral", ar.col = "red1", len = 0.2,
       xaxt = "n")
  
  # Plot B: Procrustes plot for Non-littoral
  plot(procrustes_list_all_nolitt_az$procrustes_result, kind = 1, type = "text", cex = 1, main = "Non-littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", ylab = "")

  
  dev.off()





# Kind 2


  png(here::here("c_output/figures/procrustes_kind2.png"), width = 25, height = 26, units = "cm", res = 600) 
  par(mfrow = c(3, 2))
  
  
  # Plot A: Procrustes plot for Littoral
  plot(procrustes_list_all_litt_gal$procrustes_result, kind = 2, type = "text", cex = 1, main = "Littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "")
  axis(side = 1, at = 1:nrow(procrustes_list_all_litt_gal$procrustes_result$Yrot), 
       labels = rownames(procrustes_list_all_litt_gal$procrustes_result$Yrot), tick = TRUE, las = 2)
  
  # Plot B: Procrustes plot for Non-littoral
  plot(procrustes_list_all_nolitt_gal$procrustes_result, kind = 2, type = "text", cex = 1, main = "Non-littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "", ylab = "")
    axis(side = 1, at = 1:nrow(procrustes_list_all_nolitt_gal$procrustes_result$Yrot), 
       labels = rownames(procrustes_list_all_nolitt_gal$procrustes_result$Yrot), tick = TRUE, las = 2)
  
    # Plot C: Procrustes plot for Littoral
  plot(procrustes_list_all_litt_can$procrustes_result, kind = 2, type = "text", cex = 1, main = "Littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "")
    axis(side = 1, at = 1:nrow(procrustes_list_all_litt_can$procrustes_result$Yrot), 
       labels = rownames(procrustes_list_all_litt_can$procrustes_result$Yrot), tick = TRUE, las = 2)
  
  # Plot D: Procrustes plot for Non-littoral
  plot(procrustes_list_all_nolitt_can$procrustes_result, kind = 2, type = "text", cex = 1, main = "Non-littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", xlab = "", ylab = "")
    axis(side = 1, at = 1:nrow(procrustes_list_all_nolitt_can$procrustes_result$Yrot), 
       labels = rownames(procrustes_list_all_nolitt_can$procrustes_result$Yrot), tick = TRUE, las = 2)
  
    # Plot A: Procrustes plot for Littoral
  plot(procrustes_list_all_litt_az$procrustes_result, kind = 2, type = "text", cex = 1, main = "Littoral", ar.col = "red1", len = 0.2,
       xaxt = "n")
    axis(side = 1, at = 1:nrow(procrustes_list_all_litt_az$procrustes_result$Yrot), 
       labels = rownames(procrustes_list_all_litt_az$procrustes_result$Yrot), tick = TRUE, las = 2)
  
  # Plot B: Procrustes plot for Non-littoral
  plot(procrustes_list_all_nolitt_az$procrustes_result, kind = 2, type = "text", cex = 1, main = "Non-littoral", ar.col = "red1", len = 0.2,
       xaxt = "n", ylab = "")
    axis(side = 1, at = 1:nrow(procrustes_list_all_nolitt_az$procrustes_result$Yrot), 
       labels = rownames(procrustes_list_all_nolitt_az$procrustes_result$Yrot), tick = TRUE, las = 2)

  
  dev.off()

```