---
title: "Compute littoral percent and centrality measures of islands"
---

```{r}
library(ggplot2)
library(MASS)
library(tidyverse) 
library(gdistance)
library(fields)
library(igraph)
library(tidygraph)
library(ggrepel)
library(reshape2)
library(ggpubr)
library(DHARMa)
library(car)

source("Functions.R")
```






### Read directed matrices currents

```{r}
mat_median_min_singlepoints_gal <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_gal.rds"))
mat_median_min_singlepoints_can <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_can.rds"))
mat_median_min_singlepoints_az <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_az.rds"))

```


### Read floristic data

```{r}
data_Galapagos <- read_csv(here::here("b_data/Data_Galapagos/data_galapagos.csv"), col_names = TRUE, trim_ws = TRUE)
data_Canaries <- read_csv(here::here("b_data/Data_Canaries/data_canaries.csv"), col_names = TRUE, trim_ws = TRUE)
data_Azores <- read_csv(here::here("b_data/Data_Azores/data_azores.csv"), col_names = TRUE)

df_Galapagos <- as.data.frame(data_Galapagos)
df_Galapagos <- unite(df_Galapagos, name_species, c(Genus, Species), remove=FALSE)
df_Galapagos$name_species <- gsub('Sesuvium_edmondstonei', 'Sesuvium_edmonstonei', df_Galapagos$name_species)
df_Galapagos$name_species <- gsub('Pycreus_polystachyos', 'Cyperus_polystachyos', df_Galapagos$name_species)

df_Canaries <- as.data.frame(data_Canaries)
df_Azores <- as.data.frame(data_Azores)

df_Azores <- df_Azores %>% 
  rename(
    Corvo = COR,
    Faial = FAI,
    Flores = FLO,
    Graciosa = GRA,
    Pico = PIC,
    SantaMaria = SMR,
    SaoJorge = SJG,
    SaoMiguel = SMG,
    Terceira = TER
  )
```

Filter lowland (azores database already filtered)

```{r}
df_Galapagos_lowland <- subset(df_Galapagos, df_Galapagos$habitats_lessthan500m %in% "1")
df_Canaries_lowland <- subset(df_Canaries, df_Canaries$habitats_lessthan500m %in% "1")
```


## Compute percentage littoral species per island


Galapagos

```{r}
# Define the columns for islands
island_cols <- c("Espanola", "Fernandina", "Floreana", "Genovesa", "Isabela", "Marchena", "Pinta", "Pinzon", "San_Cristobal", "Santa_Cruz", "Santa_Fe", "Santiago")

# Compute the total number of littoral species in the archipelago
total_littoral_species_archip <- nrow(df_Galapagos_lowland[df_Galapagos_lowland$littoral %in% c("exclusive", "presence"), ])

# Initialize a dataframe to store the results
results <- data.frame(
  Island = character(),
  Percent_local_floras = numeric(),
  Percent_regional_litt = numeric(),
  Total_Species = integer(),
  Total_Litt_archip = integer(),
  stringsAsFactors = FALSE
)

# Iterate over each island
for (island in island_cols) {
  # Subset the species present in the island
  species_in_island <- df_Galapagos_lowland[complete.cases(df_Galapagos_lowland[, island]), ]
  
  # Calculate the total number of species in the island
  total_species_island <- nrow(species_in_island)
  
  # Calculate the percentage of littoral species relative to all species on the island
  littoral_species_island <- species_in_island[species_in_island$littoral %in% c("exclusive", "presence"), ]
  Percent_local_floras <- (nrow(littoral_species_island) / total_species_island) * 100
  
  # Calculate the percentage of littoral species relative to the archipelago's littoral species
  Percent_regional_litt <- (nrow(littoral_species_island) / total_littoral_species_archip) * 100
  
  # Add the results to the dataframe
  results <- rbind(results, data.frame(
    Island = island,
    Percent_local_floras = Percent_local_floras,
    Percent_regional_litt = Percent_regional_litt,
    Littoral_Species_island = nrow(littoral_species_island),
    Total_Species_island = total_species_island,
    Total_Litt_archip = total_littoral_species_archip,
    stringsAsFactors = FALSE
  ))
}

df_percent_flora_gal <- results


```



Canaries

```{r}
# Define the columns for islands
island_cols <- c("El_Hierro", "La_Palma", "La_Gomera", "Tenerife", "Gran_Canaria", "Fuerteventura", "Lanzarote")

# Compute the total number of littoral species in the archipelago
total_littoral_species_archip <- nrow(df_Canaries_lowland[df_Canaries_lowland$littoral %in% c("exclusive", "presence"), ])

# Initialize a dataframe to store the results
results <- data.frame(
  Island = character(),
  Percent_local_floras = numeric(),
  Percent_regional_litt = numeric(),
  Total_Species_island = integer(),
  Littoral_Species_island = integer(),
  Total_Litt_archip = integer(),
  stringsAsFactors = FALSE
)

# Iterate over each island
for (island in island_cols) {
  # Subset the species present in the island
  species_in_island <- df_Canaries_lowland[df_Canaries_lowland[[island]] == 1, ]
  
  # Calculate the total number of species in the island
  total_species_island <- nrow(species_in_island)
  
  # Calculate the percentage of littoral species relative to all species on the island
  littoral_species_island <- species_in_island[species_in_island$littoral %in% c("exclusive", "presence"), ]
  Percent_local_floras <- (nrow(littoral_species_island) / total_species_island) * 100
  
  # Calculate the percentage of littoral species relative to the archipelago's littoral species
  Percent_regional_litt <- (nrow(littoral_species_island) / total_littoral_species_archip) * 100
  
  # Add the results to the dataframe
  results <- rbind(results, data.frame(
    Island = island,
    Percent_local_floras = Percent_local_floras,
    Percent_regional_litt = Percent_regional_litt,
    Littoral_Species_island = nrow(littoral_species_island),
    Total_Species_island = total_species_island,
    Total_Litt_archip = total_littoral_species_archip,
    stringsAsFactors = FALSE
  ))
}

df_percent_flora_can <- results



```


Azores

```{r}
# Define the columns for islands
island_cols <- c("SantaMaria", "SaoMiguel", "Terceira", "Graciosa", "SaoJorge", "Pico", "Faial", "Flores", "Corvo")

# Compute the total number of littoral species in the archipelago
total_littoral_species_archip <- nrow(df_Azores[df_Azores$littoral %in% 1, ])

# Initialize a dataframe to store the results
results <- data.frame(
  Island = character(),
  Percent_local_floras = numeric(),
  Percent_regional_litt = numeric(),
  Total_Species_island = integer(),
  Littoral_Species_island = integer(),
  Total_Litt_archip = integer(),
  stringsAsFactors = FALSE
)

# Iterate over each island
for (island in island_cols) {
  # Subset the species present in the island
  species_in_island <- df_Azores[df_Azores[[island]] == 1, ]
  
  # Calculate the total number of species in the island
  total_species_island <- nrow(species_in_island)
  
  # Calculate the percentage of littoral species relative to all species on the island
  littoral_species_island <- species_in_island[species_in_island$littoral %in% 1, ]
  Percent_local_floras <- (nrow(littoral_species_island) / total_species_island) * 100
  
  # Calculate the percentage of littoral species relative to the archipelago's littoral species
  Percent_regional_litt <- (nrow(littoral_species_island) / total_littoral_species_archip) * 100
  
  # Add the results to the dataframe
  results <- rbind(results, data.frame(
    Island = island,
    Percent_local_floras = Percent_local_floras,
    Percent_regional_litt = Percent_regional_litt,
    Littoral_Species_island = nrow(littoral_species_island),
    Total_Species_island = total_species_island,
    Total_Litt_archip = total_littoral_species_archip,
    stringsAsFactors = FALSE
  ))
}

df_percent_flora_az <- results


```


Combine data into a single table

```{r}
# Add the archipelago column to each dataframe
df_percent_flora_gal$Archipelago <- "Galapagos"
df_percent_flora_can$Archipelago <- "Canaries"
df_percent_flora_az$Archipelago <- "Azores"

# Combine the dataframes
df_percent_flora_combined <- rbind(
  df_percent_flora_gal,
  df_percent_flora_can,
  df_percent_flora_az
)

# Reorder columns to have Archipelago as the first column
df_percent_flora_combined <- df_percent_flora_combined[, c("Archipelago", names(df_percent_flora_combined)[names(df_percent_flora_combined) != "Archipelago"])]

df_percent_flora_combined$Island <- gsub("_", "", df_percent_flora_combined$Island)

df_percent_flora_combined

```



# Compute centrality measures per island

We use the inverted weights to consider the degree strenghts, as we are taking into account the sea currents network.

To compute betwennness, edges are interpreted as distances, not as connection strengths. Therefore, we need to use the true weights.

```{r}

compute_df_percent_flora_degree <- function(mat_min_conn_currents, df_percent) {
  
  g <- graph_from_adjacency_matrix(mat_min_conn_currents, mode = "directed", weighted = TRUE)
  
  # Create a data frame from the igraph object
  node_list <- data.frame(name = V(g)$name)
  
   # Merge the coordinates with the node list
  node_list <- merge(node_list, df_percent[,c("Island", "Percent_local_floras", "Percent_regional_litt")], by.x = "name", by.y = "Island", all.x = TRUE)
  
  # Ensure the graph vertices match the island names
  V(g)$name <- as.character(V(g)$name)
  node_list$name <- as.character(node_list$name)
  
  # Convert the igraph object to a tidygraph object
  tg <- as_tbl_graph(g)
  
    # Add coordinates to the tidygraph object
  tg <- tg %>%
    activate(nodes) %>%
    left_join(node_list, by = c("name" = "name"))
  
    # Invert the weights: higher values become lower and vice versa
  E(tg)$inv_weight <- 1 / E(tg)$weight
  
  
  ### CENTRALITY MEASURES
  
  # in degree (sinks)
  V(tg)$in_degree <- strength(tg, mode = "in", weights = E(tg)$inv_weight)
  
  # out_degree (pulls)
  V(tg)$out_degree <- strength(tg, mode = "out", weights = E(tg)$inv_weight)
  
    # Closeness
  V(tg)$closeness <- closeness(tg, weights = E(tg)$weight)
  
  # Betwenness
  V(tg)$betweenness <- betweenness(tg, weights = E(tg)$weight)
  

  

  df_percent_litt_degree_az <- data.frame(tg)
  
  return(df_percent_litt_degree_az)
  
}

# Galapagos

df_percent_flora_gal[which(df_percent_flora_gal$Island == "San_Cristobal"), "Island"] <- "SanCristobal"
df_percent_flora_gal[which(df_percent_flora_gal$Island == "Santa_Cruz"), "Island"] <- "SantaCruz"
df_percent_flora_gal[which(df_percent_flora_gal$Island == "Santa_Fe"), "Island"] <- "SantaFe"

df_percent_centrality_gal <- compute_df_percent_flora_degree(mat_median_min_singlepoints_gal, df_percent_flora_gal)
df_percent_centrality_gal$archipelago <- rep("Galapagos", times = nrow(df_percent_centrality_gal))


# Canaries
  
df_percent_flora_can[which(df_percent_flora_can$Island == "El_Hierro"), "Island"] <- "ElHierro"
df_percent_flora_can[which(df_percent_flora_can$Island == "Gran_Canaria"), "Island"] <- "GranCanaria"
df_percent_flora_can[which(df_percent_flora_can$Island == "La_Gomera"), "Island"] <- "LaGomera"
df_percent_flora_can[which(df_percent_flora_can$Island == "La_Palma"), "Island"] <- "LaPalma"


df_percent_centrality_can <- compute_df_percent_flora_degree(mat_median_min_singlepoints_can, df_percent_flora_can)
df_percent_centrality_can$archipelago <- rep("Canaries", times = nrow(df_percent_centrality_can))


# Azores
  
df_percent_centrality_az <-compute_df_percent_flora_degree(mat_median_min_singlepoints_az, df_percent_flora_az)
df_percent_centrality_az$archipelago <- rep("Azores", times = nrow(df_percent_centrality_az))



df_percent_centrality_all <- rbind(df_percent_centrality_gal, df_percent_centrality_can, df_percent_centrality_az)

# For readability:

#df_percent_degree_all$closeness <- df_percent_degree_all$closeness*1000

df_percent_centrality_all <- df_percent_centrality_all %>% 
  rename(island = name)

options(scipen = 999)

df_percent_centrality_all

```


Add information on total plants per island and total littoral plants per island and archipelago

```{r}
df_percent_centrality_litt_all <- merge(
  df_percent_centrality_all,
  df_percent_flora_combined[, c("Archipelago", "Island", "Littoral_Species_island", "Total_Species_island", "Total_Litt_archip")],
  by.x = c("island", "archipelago"),
  by.y = c("Island", "Archipelago")
)

df_percent_centrality_litt_all <- df_percent_centrality_litt_all %>%
  arrange(archipelago, island) %>%
  select(archipelago, everything())

df_percent_centrality_litt_all
```




# Add centrality by geographic distances


```{r}
# read matrices

mat_geo_gal <- read.csv(here::here("b_data/Data_Galapagos/geographic_distance_galap.csv"))
rownames(mat_geo_gal) <- mat_geo_gal[,1]
mat_geo_gal<-mat_geo_gal[,-1]

mat_geo_can <- read.csv(here::here("b_data/Data_Canaries/geographic_distance_canaries.csv"))
rownames(mat_geo_can) <- mat_geo_can[,1]
mat_geo_can<-mat_geo_can[,-1]

mat_geo_az <- read.csv(here::here("b_data/Data_Azores/geographic_distance_azores.csv"))
rownames(mat_geo_az) <- mat_geo_az[,1]
mat_geo_az<-mat_geo_az[,-1]


# order names and convert to matrices

mat_geo_gal <- as.matrix(mat_geo_gal[order(rownames(mat_geo_gal)), order(colnames(mat_geo_gal))])

mat_geo_can <- as.matrix(mat_geo_can[order(rownames(mat_geo_can)), order(colnames(mat_geo_can))])

mat_geo_az <- as.matrix(mat_geo_az[order(rownames(mat_geo_az)), order(colnames(mat_geo_az))])
```


```{r}
# Create graph objects for geographic distances (undirected)
create_geo_graphs <- function(geo_mat) {
  g_geo <- graph_from_adjacency_matrix(geo_mat, mode = "undirected", weighted = TRUE)
  V(g_geo)$name <- as.character(V(g_geo)$name)
  return(g_geo)
}

galapagos_geo <- create_geo_graphs(mat_geo_gal)
canaries_geo <- create_geo_graphs(mat_geo_can)
azores_geo <- create_geo_graphs(mat_geo_az)

# Function to compute centrality measures for geographic distances
compute_geo_centralities <- function(g_geo) {
  centralities <- data.frame(
    island = V(g_geo)$name,
    degree_geo = strength(g_geo, mode = "all", weights = E(g_geo)$weight),
    closeness_geo = closeness(g_geo, weights = E(g_geo)$weight),
    betweenness_geo = betweenness(g_geo, weights = E(g_geo)$weight)
  )
  return(centralities)
}

galapagos_centralities <- compute_geo_centralities(galapagos_geo)
canaries_centralities <- compute_geo_centralities(canaries_geo)
azores_centralities <- compute_geo_centralities(azores_geo)

# Add archipelago information
galapagos_centralities$archipelago <- "Galapagos"
canaries_centralities$archipelago <- "Canaries"
azores_centralities$archipelago <- "Azores"

centralities_geo <- rbind(galapagos_centralities, canaries_centralities, azores_centralities)

centralities_geo$island <- gsub("_", "", centralities_geo$island)

# Combine dataframes
df_centrality_cur_geo_litt <- df_percent_centrality_litt_all %>%
  inner_join(centralities_geo, by = c("island", "archipelago"))

df_centrality_cur_geo_litt

```


Adjustements to the dataframe


```{r}
# 1. Inverse the value of $degree_geo
df_centrality_cur_geo_litt$degree_geo <- 1 / df_centrality_cur_geo_litt$degree_geo

# 2. Multiply the values of $closeness and $closeness_geo by 1000 to ease readability
df_centrality_cur_geo_litt$closeness <- df_centrality_cur_geo_litt$closeness * 1000
df_centrality_cur_geo_litt$closeness_geo <- df_centrality_cur_geo_litt$closeness_geo * 1000

# 3. Reorder the columns
df_centrality_cur_geo_litt <- df_centrality_cur_geo_litt[, c(
  "archipelago",
  "island",
  "degree_geo",
  "in_degree",
  "out_degree",
  "closeness_geo",
  "closeness",
  "betweenness_geo",
  "betweenness",
  "Percent_local_floras",
  "Percent_regional_litt",
  "Littoral_Species_island",
  "Total_Species_island",
  "Total_Litt_archip"
)]

# View the modified dataframe
df_centrality_cur_geo_litt
```

**IMPORTANT NOTE**: the degree_geo, and in and out degree values for currents are inverted


## Add island information for using it in the models

#### Load island area matrices and convert to distance objects

Area

```{r}
# Read and process the Galapagos area data
df_area_gal <- read.csv(here::here("b_data/Data_Galapagos/area_islands_galap.csv")) %>%
  column_to_rownames('X') %>%
  as.data.frame() %>%
  mutate(archipelago = "Galapagos")

# Read and process the Canaries area data
df_area_can <- read.csv(here::here("b_data/Data_Canaries/area_islands_canaries.csv")) %>%
  column_to_rownames('X') %>%
  as.data.frame() %>%
  mutate(archipelago = "Canaries")

# Read and process the Azores area data
df_area_az <- read.csv(here::here("b_data/Data_Azores/area_islands_azores.csv")) %>%
  column_to_rownames('X') %>%
  as.data.frame() %>%
  mutate(archipelago = "Azores")

# Combine all datasets
df_area_islands <- rbind(df_area_gal, df_area_can, df_area_az)

# Convert rownames to a column called "island"
df_area_islands$island <- rownames(df_area_islands)

# Reset rownames
rownames(df_area_islands) <- NULL

# Correct names format
df_area_islands$island <- gsub("_", "", df_area_islands$island)

df_area_islands
```



Ages

```{r}
# Read and process the Galapagos data
df_age_gal <- read.csv(here::here("b_data/Data_Galapagos/ages_islands_galap.csv"), sep = ";") %>%
  column_to_rownames('island') %>%
  as.data.frame() %>%
  mutate(archipelago = "Galapagos")

# Read and process the Canaries data
df_age_can <- read.csv(here::here("b_data/Data_Canaries/ages_islands_canaries.csv"), sep = ";") %>%
  column_to_rownames('island') %>%
  as.data.frame() %>%
  mutate(archipelago = "Canaries")

# Read and process the Azores data
df_age_az <- read.csv(here::here("b_data/Data_Azores/ages_islands_azores.csv"), sep = ";") %>%
  column_to_rownames('island') %>%
  as.data.frame() %>%
  mutate(archipelago = "Azores")

# Combine all datasets
df_age_islands <- rbind(df_age_gal, df_age_can, df_age_az)

# Convert rownames to a column called "island"
df_age_islands$island <- rownames(df_age_islands)

# Reset rownames
rownames(df_age_islands) <- NULL

# correct names format
df_age_islands$island <- gsub("_", "", df_age_islands$island)

df_age_islands
```

```{r}
# Join df_centrality_cur_geo_litt with area data
df_centrality_cur_geo_litt <- df_centrality_cur_geo_litt %>%
  left_join(df_area_islands %>% select(island, archipelago, area), 
            by = c("island", "archipelago"))

# Join df_centrality_cur_geo_litt with age data
df_centrality_cur_geo_litt <- df_centrality_cur_geo_litt %>%
  left_join(df_age_islands %>% select(island, archipelago, age), 
            by = c("island", "archipelago"))

df_centrality_cur_geo_litt
```

save data

```{r}
saveRDS(df_centrality_cur_geo_litt, here::here("c_output/df_centrality_cur_geo_litt.rds"))
write.csv(df_centrality_cur_geo_litt, here::here("c_output/tables/df_centrality_cur_geo_litt.csv"))
#df_centrality_cur_geo_litt <- readRDS(here::here("c_output/df_centrality_cur_geo_litt.rds"))

df_centrality_cur_geo_litt
```



```{r}
# Define colors for each archipelago
colors <- c("Galapagos" = "Skyblue", "Canaries" = "Lightgreen", "Azores" = "Lightcoral")

my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))

```




## Plot % littoral species per island

```{r}


create_plot_floras <- function(data, archipelago_name) {
  plot_data <- data %>%
    filter(archipelago == archipelago_name) %>%
    arrange(desc(Percent_local_floras)) %>%
    mutate(island = factor(island, levels = unique(island)))  # Order islands by value
  
  y_limits <- c(0, max(plot_data$Percent_local_floras) * 1.3)
  
  ggplot(plot_data, aes(x = island, y = Percent_local_floras, fill = archipelago)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("Galapagos" = "skyblue", "Canaries" = "lightgreen", "Azores" = "lightcoral")) +
    geom_text(aes(label = sprintf("%.1f", age)), vjust = -2, size = 3.2) +
    geom_text(aes(label = sprintf("%.0f", area)), vjust = -0.5, size = 3.2) +
    ylim(y_limits) +
    theme_classic() +
    my_theme+
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),  # Angle labels for readability
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 12),
      plot.title = element_text(hjust = 0.5, size = 10),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "none"
    ) +
    labs(title = " ", y = "% littoral species")
}

archipelago_order <- c("Galapagos", "Canaries", "Azores")

# Create plots for each archipelago
plots_floras <- lapply(archipelago_order, function(archipelago) {
  create_plot_floras(df_centrality_cur_geo_litt, archipelago)
})

# Combine plots into one
plot_percent_local_floras <- ggarrange(
  plotlist = plots_floras,
  nrow = 3, ncol = 1,
  labels = LETTERS[1:3]
)

plot_percent_local_floras

ggsave(here::here("c_output/figures/plot_perclitt_islands.png"), height = 9, width = 4)

```



## Plot data geographic centrality

```{r}

df_long <- df_centrality_cur_geo_litt %>%
  pivot_longer(cols = c(degree_geo, closeness_geo, betweenness_geo), 
               names_to = "centrality", 
               values_to = "value")


create_plot <- function(data, archipelago_name, centrality_name) {
  # Define y-axis labels based on centrality_name
  y_axis_label <- switch(
    centrality_name,
    "degree_geo" = "Degree (Geographic)",
    "closeness_geo" = "Closeness (Geographic)",
    "betweenness_geo" = "Betweenness (Geographic)"
  )
  
  # Filter and order the data for the specific archipelago and centrality
  plot_data <- data %>%
    filter(archipelago == archipelago_name & centrality == centrality_name) %>%
    arrange(desc(value)) %>%
    mutate(island = factor(island, levels = unique(island)))  # Order islands by value
  
  # Define y-axis limit
  y_limit <- max(plot_data$value) * 1.3  # Increase the limit by 30% for text space
  
  ggplot(plot_data, aes(x = island, y = value, fill = archipelago)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = sprintf("%.1f", age)), 
              vjust = -2, 
              size = 3) +  # Age annotation on top
    geom_text(aes(label = sprintf("%.0f", area)), 
              vjust = -0.5, 
              size = 3) +  # Area annotation below the age
    scale_fill_manual(values = colors) +
    theme_classic() +
    my_theme +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Angle labels for readability
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 10),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "none") +
    labs(title = " ", y = y_axis_label) +
    ylim(0, y_limit)
}



# Create plots for each combination of archipelago and centrality
plots <- list()

for (archipelago in unique(df_long$archipelago)) {
  for (centrality in unique(df_long$centrality)) {
    p <- create_plot(df_long, archipelago, centrality)
    plots[[paste(archipelago, centrality, sep = "_")]] <- p
  }
}

# Arrange plots into a 3x3 grid
plot_geo_centrality_islands <- ggarrange(
  plots[["Galapagos_degree_geo"]], plots[["Galapagos_closeness_geo"]], plots[["Galapagos_betweenness_geo"]],
  plots[["Canaries_degree_geo"]], plots[["Canaries_closeness_geo"]], plots[["Canaries_betweenness_geo"]],
  plots[["Azores_degree_geo"]], plots[["Azores_closeness_geo"]], plots[["Azores_betweenness_geo"]],
  nrow = 3, ncol = 3,
  labels = LETTERS[1:9]
)

plot_geo_centrality_islands

#ggsave(here::here("c_output/figures/plot_geo_centrality_islands.png"), height = 10, width = 11)

```




## plot data ocean current centrality

```{r}


# Pivot data for centrality measures based on currents
df_long_currents <- df_centrality_cur_geo_litt %>%
  pivot_longer(cols = c(in_degree, out_degree, closeness, betweenness), 
               names_to = "centrality", 
               values_to = "value")

# Transform in_degree and out_degree to log for Canaries
df_long_currents <- df_long_currents %>%
  mutate(value = case_when(
    archipelago == "Canaries" & centrality == "in_degree" ~ log(value),
    archipelago == "Canaries" & centrality == "out_degree" ~ log(value),
    TRUE ~ value
  ))

create_plot_currents <- function(data, archipelago_name, centrality_name) {
  # Define y-axis labels based on centrality_name
  y_axis_label <- switch(
    centrality_name,
    "in_degree" = if (archipelago_name == "Canaries") "log(In-Degree) (currents)" else "In-degree (currents)",
    "out_degree" = if (archipelago_name == "Canaries") "log(Out-Degree) (currents)" else "Out-degree (currents)",
    "closeness" = "Closeness (currents)",
    "betweenness" = "Betweenness (currents)"
  )
  
  # Filter and order the data for the specific archipelago and centrality
  plot_data <- data %>%
    filter(archipelago == archipelago_name & centrality == centrality_name) %>%
    arrange(desc(value)) %>%
    mutate(island = factor(island, levels = unique(island)))  # Order islands by value
  
  # Define y-axis limits dynamically for Canaries in-degree and out-degree
  y_limits <- if (archipelago_name == "Canaries" & centrality_name %in% c("in_degree", "out_degree")) {
    c(min(plot_data$value) * 1.5, max(plot_data$value) * 4)
  } else {
    c(0, max(plot_data$value) * 1.3)
  }
  
  ggplot(plot_data, aes(x = island, y = value, fill = archipelago)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = sprintf("%.1f", age)), 
              vjust = ifelse(plot_data$value >= 0, -2, -2), 
              size = 3) +  # Age annotation on top
    geom_text(aes(label = sprintf("%.0f", area)), 
              vjust = ifelse(plot_data$value >= 0, -0.5, -0.5), 
              size = 3) +  # Area annotation below the age
    scale_fill_manual(values = colors) +
    theme_classic() +
    my_theme +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Angle labels for readability
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 10),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "none") +
    labs(title = " ", y = y_axis_label) +
    ylim(y_limits)
}


# Create plots for each combination of archipelago and centrality
plots_currents <- list()

for (archipelago in unique(df_long_currents$archipelago)) {
  for (centrality in unique(df_long_currents$centrality)) {
    p <- create_plot_currents(df_long_currents, archipelago, centrality)
    plots_currents[[paste(archipelago, centrality, sep = "_")]] <- p
  }
}

# Arrange plots into a 3x4 grid
plot_currents_centrality_islands <- ggarrange(
  plots_currents[["Galapagos_in_degree"]], plots_currents[["Galapagos_out_degree"]], 
  plots_currents[["Galapagos_closeness"]],
  plots_currents[["Galapagos_betweenness"]],
  plots_currents[["Canaries_in_degree"]],
  plots_currents[["Canaries_out_degree"]], 
  plots_currents[["Canaries_closeness"]],
  plots_currents[["Canaries_betweenness"]],
  plots_currents[["Azores_in_degree"]], 
  plots_currents[["Azores_out_degree"]], 
  plots_currents[["Azores_closeness"]],
  plots_currents[["Azores_betweenness"]],
  nrow = 3, ncol = 4,
  labels = LETTERS[1:12]
)

plot_currents_centrality_islands

#ggsave(here::here("c_output/figures/plot_currents_centrality_islands.png"), height = 10, width = 14)

```


