---
title: "Function get.ycom"
output: html_notebook
---


# Script Matrix_Con_Currs_HYCOM.Rmd


The funcion builds a character string pointing to the url where certain variables (currents u and v for a certain day) are stored. This is done by concatenating the data for the day with the common parts of the url that are fixed between different urls. Data on the url is downloaded at the end of the script with the function 'curl_download()', and stored as a netcdf file on disk.

```{r}

get.hycom <- function(limits, time, vars=c('water_u', 'water_v'), include_latlon=TRUE,
                          filename='', download.file=TRUE, dir = getwd(), depLevels=NULL) {
  
  ## Set the base URL based on the start date. If the ending date exceeds the
  ## period for this experiment, then print a warning and truncate the output
  ## early.
  expts = data.frame(
    start=c(as.Date('1992-10-02'), as.Date('1995-08-01'),
            as.Date('2013-01-01'), as.Date('2013-08-21'),
            as.Date('2014-04-05'), as.Date('2016-04-18')),
    end=c(as.Date('1995-07-31'), as.Date('2012-12-31'),
          as.Date('2013-08-20'), as.Date('2014-04-04'),
          as.Date('2016-04-17'), Sys.Date() + 1),
    url=c('http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_19.0/',
          'http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_19.1/',
          'http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_90.9?',
          'http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_91.0?',
          'http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_91.1?',
          'http://ncss.hycom.org/thredds/ncss/GLBu0.08/expt_91.2?'))
  
  if(time[1] < expts$start[1])
    stop('Data begins at %s and is not available at %s.',
         strftime(expts$start[1], '%d %b %Y'),
         strftime(time[1], '%d %b %Y'))
  if(time[1] > expts$end[nrow(expts)])
    stop('Data ends at %s and is not available at %s.',
         strftime(expts$end[nrow(expts)], '%d %b %Y'),
         strftime(time[1], '%d %b %Y'))
  for(i in seq(nrow(expts))) {
    if((time[1] >= expts$start[i]) & (time[1] <= expts$end[i]))
      url = expts$url[i]
  }
  
  if(any(grep('19', url))) url = sprintf('%s%s?', url, as.numeric(format(time, '%Y')))
  
  
  
  ## Add the variables.
  for(var in vars)
    url = sprintf('%svar=%s&', url, var)
  
  
  
  ## Add the spatial domain.
  url = sprintf('%snorth=%f&west=%f&east=%f&south=%f&horizStride=1&',
                url, limits[[4]], limits[[1]], limits[[2]], limits[[3]])
  # north, west, east, south
  
  
  
  ## Add the time domain.
  if(length(time) == 2){
    url = sprintf('%stime_start=%s%%3A00%%3A00Z&time_end=%s%%3A00%%3A00Z&timeStride=1&',
                  url, strftime(time[1], '%Y-%m-%dT00'),
                  strftime(time[2], '%Y-%m-%dT00'))
  } else if(length(time) == 1){
    url = sprintf('%stime_start=%s%%3A00%%3A00Z&time_end=%s%%3A00%%3A00Z&timeStride=1&',
                  url, strftime(time[1], '%Y-%m-%dT00'),
                  strftime(time[1], '%Y-%m-%dT00'))
  }
  
  
  
  ## Add the lat-lon points if requested.
  if(include_latlon)
    url = sprintf('%saddLatLon=true&', url)
  
  
  
  ## Finish the URL.
  if (is.null(depLevels)){
    url = sprintf('%sdisableProjSubset=on&vertCoord=&accept=netcdf', url)
  } else{
    url = paste(url,'disableProjSubset=on&vertCoord=', depLevels, '&accept=netcdf', sep='')
  }
  
  print(url)

  
  
  ## Download the data if a filename was provided.
  if(filename != ''){
    if(download.file == TRUE){
      #download.file(url, filename, method = 'auto')
      curl_download(url, filename, quiet=FALSE)
    } else if(download.file == FALSE){
      system(sprintf('curl -o "%s" "%s"', filename, url))
    }
  }
  return(url)
}
```







```{r}

calculate_costMatrix <- function(files, geo_points) {
  
  
  cost_list <- list()
  
  for(i in 1:length(files)){
  
  
  
  cat("File ", i, "/n")  # to keep track of analyzed files
  
  #Read u and v from the netcdf raster (.nc extension)
  
  # nc_open from the netCDF package allows for getting the names of bands on 
  #other specific properties of netCDF that are lacking in the raster package
  
  nc <- nc_open(files[i])
  if(length(which(abs(nc$var$water_u$dim[[1]]$vals) > 180)) >= 1){  #depending on their coords, some rasters need to be rotated
    rr_u <- rotate(raster(files[i], level = 1, varname = "water_u", band = 1))
    rr_v <- rotate(raster(files[i], level = 1, varname = "water_v", band = 1))
    
  } else {
    rr_u <- raster(files[i], level = 1, varname = "water_u", band = 1)
    
    rr_v <- raster(files[i], level = 1, varname = "water_v", band = 1)
    
  }
  
  # We already have currents u (rr_u) and v (rr_v). Optionally, we
  # can save these in a stack with:
  # stack_u <- stack(stack_u, rr_u)
  # stack_v <- stack(stack_v, rr_v)
  
  # Use rr_u and rr_v to calculate direction and speed
  
  
  sea_dir <- atan2(rr_u, rr_v) #from u and v components, calculatess direction
  rad2deg <- function(rad) {(rad * 180) / (pi)}
  sea_dir <- rad2deg(sea_dir)
  sea_dir[sea_dir < 0] <- 360 + sea_dir[sea_dir < 0]
  names(sea_dir) <- "direction"
  sea_spe <- sqrt( (rr_u * rr_u) + (rr_v * rr_v)) #and speed
  names(sea_spe) <- "speed"
  
  # LetÂ´s make a stack with direction and speed, that is what rWind needs to calculate conductance
  day_stack <- stack(sea_dir, sea_spe)
  
  # We now use rWind to calculate the Conductance matrix
  conductance <- flow.dispersion(day_stack, type = "passive", output = "transitionLayer")
  
  # LetÂ´s save the conductance matrix of this day in a list. IÂ´ll use the date as the name of the element within the list (each position in a list can get a name)
  #conduct_list[[i]] <- conductance
  #names(conduct_list)[i] <- files[i] %>% str_replace("_.nc", "")
  
  # And letÂ´s put the day_stack in the list
  #day_stacks_list[[i]] <- day_stack
  #names(day_stacks_list)[i] <- files[i] %>% str_replace("_.nc", "")
  
  cost_table <- costDistance(conductance, as.matrix(geo_points)) %>%    # this is the function that does it
    as_tibble %>%     # instead of a matrix, I save it as a table
    set_names(rownames(geo_points))    # with corrent column names
  
  cost_list[[i]] <- cost_table   # move the result to the list
  names(cost_list)[i] <- files[i] %>% str_replace("_.nc", "")    # and give it a proper name
  
  
  nc_close(nc)
  
  
}

return(cost_list)
  
}



```




```{r}
calculate_conductance <- function(files) {
  
  conduct_list <- list()
  cost_list <- list()
  
  for(i in 1:length(files)){
  
  
  
  cat("File ", i, "/n")  # to keep track of analyzed files
  
  #Read u and v from the netcdf raster (.nc extension)
  
  # nc_open from the netCDF package allows for getting the names of bands on 
  #other specific properties of netCDF that are lacking in the raster package
  
  nc <- nc_open(files[i])
  if(length(which(abs(nc$var$water_u$dim[[1]]$vals) > 180)) >= 1){  #depending on their coords, some rasters need to be rotated
    rr_u <- rotate(raster(files[i], level = 1, varname = "water_u", band = 1))
    rr_v <- rotate(raster(files[i], level = 1, varname = "water_v", band = 1))
    
  } else {
    rr_u <- raster(files[i], level = 1, varname = "water_u", band = 1)
    
    rr_v <- raster(files[i], level = 1, varname = "water_v", band = 1)
    
  }
  
  # We already have currents u (rr_u) and v (rr_v). Optionally, we
  # can save these in a stack with:
  # stack_u <- stack(stack_u, rr_u)
  # stack_v <- stack(stack_v, rr_v)
  
  # Use rr_u and rr_v to calculate direction and speed
  
  
  sea_dir <- atan2(rr_u, rr_v) #from u and v components, calculatess direction
  rad2deg <- function(rad) {(rad * 180) / (pi)}
  sea_dir <- rad2deg(sea_dir)
  sea_dir[sea_dir < 0] <- 360 + sea_dir[sea_dir < 0]
  names(sea_dir) <- "direction"
  sea_spe <- sqrt( (rr_u * rr_u) + (rr_v * rr_v)) #and speed
  names(sea_spe) <- "speed"
  
  # LetÂ´s make a stack with direction and speed, that is what rWind needs to calculate conductance
  day_stack <- stack(sea_dir, sea_spe)
  
  # We now use rWind to calculate the Conductance matrix
  conductance <- flow.dispersion(day_stack, type = "passive", output = "transitionLayer")
  
  # LetÂ´s save the conductance matrix of this day in a list. IÂ´ll use the date as the name of the element within the list (each position in a list can get a name)
  
  conduct_list[[i]] <- conductance
  names(conduct_list)[i] <- files[i] %>% str_replace("_.nc", "")
  
  # And letÂ´s put the day_stack in the list
  #day_stacks_list[[i]] <- day_stack
  #names(day_stacks_list)[i] <- files[i] %>% str_replace("_.nc", "")
  
  cost_table <- costDistance(conductance, as.matrix(Galapagos_points)) %>%    # this is the function that does it
    as_tibble %>%     # instead of a matrix, I save it as a table
    set_names(rownames(Galapagos_points))    # with corrent column names
  
  cost_list[[i]] <- cost_table   # move the result to the list
  names(cost_list)[i] <- files[i] %>% str_replace("_.nc", "")    # and give it a proper name
  
  
  nc_close(nc)
  
  
}

return(conduct_list)
  
}
```






Same function but return day stack list

```{r}
obtain_daystacklist <- function(files) {
  
  day_stacks_list<- list()
  conduct_list <- list()
  cost_list <- list()
  
  for(i in 1:length(files)){
  
  
  
  cat("File ", i, "/n")  # to keep track of analyzed files
  
  #Read u and v from the netcdf raster (.nc extension)
  
  # nc_open from the netCDF package allows for getting the names of bands on 
  #other specific properties of netCDF that are lacking in the raster package
  
  nc <- nc_open(files[i])
  if(length(which(abs(nc$var$water_u$dim[[1]]$vals) > 180)) >= 1){  #depending on their coords, some rasters need to be rotated
    rr_u <- rotate(raster(files[i], level = 1, varname = "water_u", band = 1))
    rr_v <- rotate(raster(files[i], level = 1, varname = "water_v", band = 1))
    
  } else {
    rr_u <- raster(files[i], level = 1, varname = "water_u", band = 1)
    
    rr_v <- raster(files[i], level = 1, varname = "water_v", band = 1)
    
  }
  
  # We already have currents u (rr_u) and v (rr_v). Optionally, we
  # can save these in a stack with:
  # stack_u <- stack(stack_u, rr_u)
  # stack_v <- stack(stack_v, rr_v)
  
  # Use rr_u and rr_v to calculate direction and speed
  
  
  sea_dir <- atan2(rr_u, rr_v) #from u and v components, calculatess direction
  rad2deg <- function(rad) {(rad * 180) / (pi)}
  sea_dir <- rad2deg(sea_dir)
  sea_dir[sea_dir < 0] <- 360 + sea_dir[sea_dir < 0]
  names(sea_dir) <- "direction"
  sea_spe <- sqrt( (rr_u * rr_u) + (rr_v * rr_v)) #and speed
  names(sea_spe) <- "speed"
  
  # LetÂ´s make a stack with direction and speed, that is what rWind needs to calculate conductance
  day_stack <- stack(sea_dir, sea_spe)
  
  # We now use rWind to calculate the Conductance matrix
  conductance <- flow.dispersion(day_stack, type = "passive", output = "transitionLayer")
  
  # LetÂ´s save the conductance matrix of this day in a list. IÂ´ll use the date as the name of the element within the list (each position in a list can get a name)
  
  #conduct_list[[i]] <- conductance
  #names(conduct_list)[i] <- files[i] %>% str_replace("_.nc", "")
  
  # And letÂ´s put the day_stack in the list
  day_stacks_list[[i]] <- day_stack
  names(day_stacks_list)[i] <- files[i] %>% str_replace("_.nc", "")
  
  cost_table <- costDistance(conductance, as.matrix(Galapagos_points)) %>%    # this is the function that does it
    as_tibble %>%     # instead of a matrix, I save it as a table
    set_names(rownames(Galapagos_points))    # with corrent column names
  
  cost_list[[i]] <- cost_table   # move the result to the list
  names(cost_list)[i] <- files[i] %>% str_replace("_.nc", "")    # and give it a proper name
  
  
  nc_close(nc)
  
  
}

return(day_stacks_list)
  
}
```




Calculate p-value matrix from shapiro test matrix

```{r}
Calculate_Pvalue_matrix <- function(shapiro_matrix) {
  
  n_value <- rep(0, 1) # value that will include the objects '[i]' from each Shapiro Test
  
  pvalue_matrix <- matrix(nrow=46, ncol=46)

  
  for (i in 1:46) {
    for (j in 1:46) {
      n_value[i] <- shapiro_matrix[i,j]
      pvalue_matrix[i,j] <- n_value[[i]]$p.value
      
    }
  }
  
 return(pvalue_matrix)
  
}
```







Calculate percentage of data distributed normally from matrix normality 

```{r}
Calculate_percent_norm_from_normMatrix <- function(normality_matrix) {
  Table_Norm_test <- table(normality_matrix) / length(normality_matrix)
  percentage_normallity <- percent(Table_Norm_test[1], accuracy = 0.01)
  return(percentage_normallity)
}


normality_matrix_Canaries



Calculate_percent_norm_from_normMatrix <- function(normality_matrix) {
  Table_Norm_test <- table(normality_matrix_Canaries) / length(normality_matrix_Canaries)
  percentage_normallity <- percent(Table_Norm_test[1], accuracy = 0.01)
  return(percentage_normallity)
}
```


Calculate matrix with the median of the 5% minimum values

```{r}
Calculate_median_minimum5percent <- function(array) {
  
  median_minimums_matrix <- matrix(nrow=46, ncol=46) # matrix that will include the median of the minimum values
  
  for (i in 1:46) {
    for (j in 1:46) {
      sorted_values <- sort(array[i,j,]) # the values from each column and row are sorted in ascending order
      x <- round(length(sorted_values)*0.05,0) # 'x' indicates how many values are the 5% of the total number of values (excluding NAs values)
      median_minimums_matrix[i,j] <- median(sorted_values[1:x]) # taken 5% of the values and done the median of these values
    } }
  
  return(median_minimums_matrix)
}
```


Calculate correlation min_matrix and median_min matrix

```{r}
Correlation_minMatrix_medianMinMatrix <- function(median_minimum_matrix, min_matrix) {
  
  r <- cor(c(min_matrix), c(median_minimums_matrix))
  R2 <- r^2
  
  return(R2)
}
```

Make matrix median min symmetric

```{r}
Make_matrix_medianMin_symmetric <- function(median_min_mat) {
  idx<-median_min_mat<=t(median_min_mat) # matrix TRUE/FALSE 
#if it's for the minimum

median_min_mat.2<-median_min_mat*idx+t(median_min_mat)*(1L-idx)

return(median_min_mat.2)

}
```


```{r}

Make_minMatrix_symmetric <- function(min_matrix) {
  idx1<-min_matrix<=t(min_matrix) # matrix TRUE/FALSE if it's for the minimum
  min_matrix.2<-min_matrix*idx1+t(min_matrix)*(1L-idx1)
  
  return(min_matrix.2)
}

```








# Script Create_Floristic_Distance_Matrices.Rmd

```{r}
# Transpose and smooth matrix, calculate Bray curtis distances and save matrix


Calculate_distmatrix_and_save <- function(matrix, file_path) {
  matrix <- t(matrix) # Transpose rows for columns
  rownames(matrix) <- c() # Delete names of the rows
  matrix[1, 1] <- NA # Delete the name of the first row
  
  colnames(matrix) <- as.character(unlist(matrix[1,])) # Put first row as header
  matrix = matrix[-1, ] # delete the first row
  
  rownames(matrix) <- matrix[,1] # Put first column as 'row header'
  matrix <- matrix[order(as.numeric(row.names(matrix)), decreasing=FALSE),] # Put first column as 'row header'
  matrix <- matrix[,-1]  # Delete the first column
  
  matrix_df <- as.data.frame(matrix) # Transform to 'data.frame'
  
  matrix_df[] <- lapply(matrix_df, function(x) as.numeric(as.character(x))) # all columns as numeric
  
  floristic_distance_bray <- vegdist(x = matrix_df, method="bray", binary=TRUE, diag=FALSE, upper=FALSE, na.rm = FALSE) 
  
  write.table(as.matrix(floristic_distance_bray), file = file_path,
              append = FALSE, quote = TRUE, sep = " ",
              eol = "\n", na = "NA", dec = ".", row.names = TRUE,
              col.names = TRUE, qmethod = c("escape", "double"),
              fileEncoding = "")
  
}
```








# Script Correlations.Rmd



```{r}
load_order_convert_table_to_distanceMat <- function(filepath) {
  flor_matrix<-read.table(filepath)
  
  
  # Order dataframe's rows and columns
flor_matrix1 <- flor_matrix[order(rownames(flor_matrix)), order(colnames(flor_matrix))]


 # convert to distance matrix
distFlor<-as.dist(flor_matrix1)

return(distFlor)

}

```



```{r}

load_order_convert_csv_to_distanceMat <- function(filepath) {
  
  # Load
  curr_matrix<-read.csv(filepath)
  
  # rownames
  rownames(curr_matrix) <- curr_matrix[,1]
curr_matrix1<-curr_matrix[,-1]

 # Order 2


curr_matrix2 <- curr_matrix1[order(rownames(curr_matrix1)), order(colnames(curr_matrix1))]

  # Convert to distance matrix
distCurr<-as.dist(curr_matrix2)


return(distCurr)
  
}
```



```{r}
Compute_Protest_distMat <- function(distMat_rot, distMat_base, k) {
  add1 <-  !(is.euclid(distMat_base))
  test1<-cmdscale(distMat_base, k = k, eig = TRUE, add = add1)
  ord1<-ordiplot(test1, type = "text", main = "PCoA Bray Distances")
  
  
  add2 <-  !(is.euclid(distMat_rot))
  test2<-cmdscale(distMat_rot, k = k, eig = TRUE, add = add2)
  ord2<-ordiplot(test2, type = "text", main = "PCoA Bray Distances")

protest <- protest(ord1,ord2,  symmetric = TRUE, scores = "sites", permutations=100000)

return(protest)
}

```



# Script IslandDegree_Nplants.Rmd


Obtain a vector with the number of plants per island. The input is a filtered floristic dataframe. 

```{r}

header.true <- function(df) {
  names(df) <- as.character(unlist(df[1,]))
  df[-1,]
}



Calculate_Nplants_perIsland <- function(df_flora) {
  df_header_names<-header.true(df_flora)
  
  island.count_sums<-colSums(df_header_names != 0)
  
  names(island.count_sums) <- NULL
  
  island.count_sums1<-island.count_sums[-1]
  
  return(island.count_sums1)
}


```


Obtain a vector with the percentage of plants per island. The input is a filtered floristic dataframe. 

```{r}

header.true <- function(df) {
  names(df) <- as.character(unlist(df[1,]))
  df[-1,]
}



fun_percentage <- function(i) {
    (i)/n_plants_total*100
}




Calculate_Percentage_plants_perIsland <- function(df_flora) {
  
  df_header_names<-header.true(df_flora)
  
  island.count_sums<-colSums(df_header_names != 0)
  
  island.count_sums1<-island.count_sums[-1] # eliminate first value which is the total counts
  
  island.count_sums1[order(names(island.count_sums1))] #order islands alphabetically
  
  names(island.count_sums1) <- NULL
  
  percentage_plants<-sapply(island.count_sums1, fun_percentage)
  
  percentage_plants_perIsland<-round(percentage_plants, digits=2)
  
  return(percentage_plants_perIsland)
  
}
```


```{r}
calculate_degrees_perIsland <- function(dist_matrix) {
  
  sim_matrix <- 1/dist_matrix # inverse values to obtain similarity matrix
  
  links_conn_curr<-dist2list(sim_matrix) # obtain edge list
  
  net_curr <- graph_from_data_frame(d=links_conn_curr, directed=F) #create graph from edgelist
  
  net_curr1 <- set_edge_attr(net_curr, "weight", value= links_conn_curr$value) # assign weight attribute
  
  degrees<-strength(
  net_curr1,
  vids = V(net_curr),
  loops = FALSE,
  weights = net_curr$value
)
  
  degrees[order(names(degrees))]
  
  names(degrees) <- NULL
  
  return(degrees)
}
```


```{r}
Create_dataframe_IslName_perc_N_degree <- function(Archipelago_name_vector,island_names_vector, Percentage_plants_vector1, Percentage_plants_vector2, Nplants_vector1, Nplants_vector2,Degree_vector1, Degree_vector2, area_vector, age_vector) {
  
  df_names_perc_N_deg<-data.frame(Archipelago_name_vector,island_names_vector, Percentage_plants_vector1, Percentage_plants_vector2, Nplants_vector1, Nplants_vector2,Degree_vector1, Degree_vector2, area_vector, age_vector)
  
  df_names_perc_N_deg1 = rename(df_names_perc_N_deg,
                                Archipelago = Archipelago_name_vector,
                                      island_names = island_names_vector,
                                      Percentage_plants_tha = Percentage_plants_vector1,
                                Percentage_plants_litt = Percentage_plants_vector2,
                                      Nplants_tha = Nplants_vector1,
                                Nplants_litt = Nplants_vector2,
                                      Degree_curr = Degree_vector1,
                                Degree_geo = Degree_vector2,
                                area = area_vector,
                                age = age_vector)
  
  return(df_names_perc_N_deg1)
  
}
```
















