---
title: "Compute network plots of current connectivities"
---



```{r}
library(ggplot2)
library(MASS)
library(tidyverse) 
library(gdistance)
library(fields)
library(scales)
library(igraph)
library(ggraph)
library(tidygraph)
library(ggrepel)
library(reshape2)
library(GGally)
library(ggpubr)
library(DHARMa)
library(car)
library(glmmTMB)

source("Functions.R")
```


parameters for plotting

```{r}
colors <- c("Galapagos" = "skyblue", "Canaries" = "#660099", "Azores" = "lightcoral")

my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))
```


Read data


Read directed matrices

```{r}
mat_median_min_singlepoints_gal <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_gal.rds"))
mat_median_min_singlepoints_can <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_can.rds"))
mat_median_min_singlepoints_az <- readRDS(here::here("c_output/current_conn_matrices/mat_median_min_singlepoints_az.rds"))

```


Read undirected matrices

```{r}

mat_medianMin_sym_Galapagos <- readRDS(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Galapagos.rds"))
mat_medianMin_sym_Canaries <- readRDS(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Canaries.rds"))
mat_medianMin_sym_Azores <- readRDS(here::here("c_output/current_conn_matrices/mat_medianMin_sym_Azores.rds"))
```



```{r}
df_centrality_cur_geo_litt <- readRDS(here::here("c_output/df_centrality_cur_geo_litt.rds"))

df_centrality_cur_geo_litt
```



## Plot networks



## Plot correlations % littoral and centrality measures



```{r}
coord_centroids_gal <- read.csv(here::here("b_data/Data_Galapagos/centroids_network_plot.csv"))
coord_centroids_can <- read.csv(here::here("b_data/Data_Canaries/centroids_network_plot.csv"))
coord_centroids_az <- read.csv(here::here("b_data/Data_Azores/centroids_network_plot.csv"))
```



```{r}
back_transform_log <- function(x) {
  return(exp(x))  # Back transform using exponential function (inverse of log)
}

```


```{r}
# Define the mathematical symbols for labels
in_degree_label <- expression(log(C[D-In]^W))
closeness_label <- expression(C[C])
```



% Littoral & centrality

```{r}

# maximum %

max_perc_litt <- max(df_centrality_cur_geo_litt$Percent_local_floras)


# Galapagos 

plot_litt_degree_gal <- ggplot(df_centrality_cur_geo_litt[df_centrality_cur_geo_litt$archipelago == "Galapagos",], aes(x = log(in_degree), y = Percent_local_floras)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
  labs(x = in_degree_label, y = "% littoral plants") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


plot_litt_cc_gal <- ggplot(df_centrality_cur_geo_litt[df_centrality_cur_geo_litt$archipelago == "Galapagos",], aes(x = closeness, y = Percent_local_floras)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
      scale_x_continuous(labels = function(x) sprintf("%.2f", back_transform_log(x))) +
  labs(x = closeness_label, y = "% littoral plants") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


# Canaries 

plot_litt_degree_can <- ggplot(df_centrality_cur_geo_litt[df_centrality_cur_geo_litt$archipelago == "Canaries",], aes(x = log(in_degree), y = Percent_local_floras)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
   scale_x_continuous(labels = function(x) sprintf("%.2f", back_transform_log(x))) +
  labs(x = in_degree_label, y = "% littoral plants") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


plot_litt_cc_can <- ggplot(df_centrality_cur_geo_litt[df_centrality_cur_geo_litt$archipelago == "Canaries",], aes(x = closeness, y = Percent_local_floras)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
      scale_x_continuous(labels = function(x) sprintf("%.3f", back_transform_log(x))) +
  labs(x = closeness_label, y = "% littoral plants") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


# Azores 

plot_litt_degree_az <- ggplot(df_centrality_cur_geo_litt[df_centrality_cur_geo_litt$archipelago == "Azores",], aes(x = log(in_degree), y = Percent_local_floras)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
  labs(x = in_degree_label, y = "% littoral plants") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme


plot_litt_cc_az <- ggplot(df_centrality_cur_geo_litt[df_centrality_cur_geo_litt$archipelago == "Azores",], aes(x = closeness, y = Percent_local_floras)) +
  geom_point(size = 5, fill = "gray40", colour="white",pch=21, size=2) + # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Add trend lines
      scale_x_continuous(labels = function(x) sprintf("%.2f", back_transform_log(x))) +
  labs(x = closeness_label, y = "% littoral plants") +
  ylim(0,max_perc_litt+1)+
  theme_classic() + # Use a minimal theme
  my_theme




plot_perc_litt_centrality <- ggarrange(
  plot_litt_degree_gal+ labs(title = ""), plot_litt_degree_can + ylab(NULL)+ labs(title = ""), plot_litt_degree_az + ylab(NULL)+ labs(title = ""),
  plot_litt_cc_gal+ labs(title = ""), plot_litt_cc_can + ylab(NULL)+ labs(title = ""), plot_litt_cc_az + ylab(NULL)+ labs(title = ""),
  
  ncol = 3,
  nrow = 2,
  
  labels = LETTERS[1:6]
)


plot_perc_litt_centrality

#ggsave(here::here("c_output/figures/plot_perc_litt_centrality.png"), height = 9, width = 9)

```




# NETWORKS


function to plot the % litt in undirected networks




## Undirected networks and % littoral flora


Currents

```{r}

df_azores <- df_centrality_cur_geo_litt |> 
  filter(archipelago == "Azores") 

df_canaries <- df_centrality_cur_geo_litt |> 
  filter(archipelago == "Canaries")

df_galapagos <- df_centrality_cur_geo_litt |> 
  filter(archipelago == "Galapagos")


# Galapagos


plot_network_litt_gal <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Galapagos,
                      centroids = coord_centroids_gal,
                      node_size = 8,
                      percent_littoral = df_galapagos[,c("Percent_local_floras", "island")],
                      scale = "none")





# Canaries

plot_network_litt_can <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Canaries,
                      centroids = coord_centroids_can,
                      node_size = 8,
                      percent_littoral = df_canaries[,c("Percent_local_floras", "island")],
                      scale = "log_scale")






# Azores


plot_network_litt_az <- plot_network_littoral(weighted_matrix = mat_medianMin_sym_Azores,
                      centroids = coord_centroids_az,
                      node_size = 8,
                      percent_littoral = df_azores[,c("Percent_local_floras", "island")],
                      scale = "sqrt_scale")


```


Plot combined network littoral

```{r}


plot_network_litt_all <- ggarrange(
  plot_network_litt_gal + xlab(""),
  plot_network_litt_can + xlab(""),
  plot_network_litt_az,
  
  labels = LETTERS[1:3],
  
  ncol = 1,
  nrow = 3
)

plot_network_litt_all

#ggsave(here::here("c_output/figures/plot_network_litt_all.png"), height = 12, width = 8)



```




## Directed networks & centrality





Plots degree

```{r}

# Galapagos

plot_network_degree_gal <- plot_network_in_degree(weighted_matrix = mat_medianMin_sym_Galapagos, 
                       centroids = coord_centroids_gal, 
                       df_degree = df_galapagos[,c("in_degree", "island")],
                       node_size = 8,
                      scale_edge = "none",
                      scale_degree = "none",
                      legend_label = in_degree_label)

# Canaries

plot_network_degree_can <- plot_network_in_degree(weighted_matrix = mat_medianMin_sym_Canaries, 
                       centroids = coord_centroids_can, 
                       df_degree = df_canaries[,c("in_degree", "island")],
                       node_size = 8,
                      scale_edge = "log_scale",
                      scale_degree = "log_scale",
                      legend_label = in_degree_label)+
    scale_color_viridis(
    name = "In-degree centrality", 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))

# Azores


plot_network_degree_az <- plot_network_in_degree(weighted_matrix = mat_medianMin_sym_Azores, 
                       centroids = coord_centroids_az, 
                       df_degree = df_azores[,c("in_degree", "island")],
                       node_size = 8,
                      scale_edge = "sqrt_scale",
                      scale_degree = "sqrt_scale",
                      legend_label = in_degree_label)

```




Plots closeness

```{r}

# Galapagos

plot_network_closeness_gal <- plot_network_closeness(weighted_matrix = mat_medianMin_sym_Galapagos, 
                       centroids = coord_centroids_gal, 
                       df_closeness = df_galapagos[,c("closeness", "island")],
                       node_size = 8,
                      scale_edge = "none",
                      scale_degree = "none",
                      legend_label = closeness_label)

# Canaries

plot_network_closeness_can <- plot_network_closeness(weighted_matrix = mat_medianMin_sym_Canaries, 
                       centroids = coord_centroids_can, 
                       df_closeness = df_canaries[,c("closeness", "island")],
                       node_size = 8,
                      scale_edge = "none",
                      scale_degree = "none",
                      legend_label = closeness_label)

# Azores


plot_network_closeness_az <- plot_network_closeness(weighted_matrix = mat_medianMin_sym_Azores, 
                       centroids = coord_centroids_az, 
                       df_closeness = df_azores[,c("closeness", "island")],
                       node_size = 8,
                      scale_edge = "none",
                      scale_degree = "none",
                      legend_label = closeness_label)

```





Plot combined network & regression degree

```{r}


plot_regr_centrality_gal <- ggarrange(
  
  plot_litt_degree_gal + ylab(NULL), NULL,
plot_litt_cc_gal+ ylab(NULL), NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_gal <- annotate_figure(
  plot_regr_centrality_gal,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


plot_regr_centrality_can <- ggarrange(
  
  plot_litt_degree_can+ ylab(NULL), NULL,
plot_litt_cc_can+ ylab(NULL), NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_can <- annotate_figure(
  plot_regr_centrality_can,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


plot_regr_centrality_az <- ggarrange(
  
  plot_litt_degree_az+ ylab(NULL), NULL,
plot_litt_cc_az+ ylab(NULL),NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_az <- annotate_figure(
  plot_regr_centrality_az,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


#############################




plot_network_regr_degree <- ggarrange(
  
      plot_network_degree_gal + theme(legend.position = "left") +
    scale_color_viridis(name = in_degree_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) 
  , plot_regr_centrality_gal,
  
    plot_network_degree_can + theme(legend.position = "left") +
    scale_color_viridis(
    name = in_degree_label, 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) 
  , plot_regr_centrality_can,
  
  plot_network_degree_az + theme(legend.position = "left") +
    scale_color_viridis(name =in_degree_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) 
  , plot_regr_centrality_az,
  
  ncol = 2,
  nrow = 3,
  
  labels = LETTERS[1:6],
  
  widths = c(2, 1)
)

plot_network_regr_degree

#ggsave(here::here("c_output/figures/plot_network_regr_degree.png"), height = 12, width = 12)


```




Plot combined network & regression closeness

```{r}


plot_regr_centrality_gal <- ggarrange(
  
  plot_litt_degree_gal + ylab(NULL), NULL,
plot_litt_cc_gal+ ylab(NULL), NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_gal <- annotate_figure(
  plot_regr_centrality_gal,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


plot_regr_centrality_can <- ggarrange(
  
  plot_litt_degree_can+ ylab(NULL), NULL,
plot_litt_cc_can+ ylab(NULL), NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_can <- annotate_figure(
  plot_regr_centrality_can,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


plot_regr_centrality_az <- ggarrange(
  
  plot_litt_degree_az+ ylab(NULL), NULL,
plot_litt_cc_az+ ylab(NULL),NULL,
nrow = 2,
ncol = 2,

widths = c(1.6, 1)
)

plot_regr_centrality_az <- annotate_figure(
  plot_regr_centrality_az,
  left = text_grob("% littoral species", rot = 90, vjust = 0.5, hjust = 0.5, size = 16)
)


#############################



plot_network_regr_closeness <- ggarrange(
  
         plot_network_closeness_gal + theme(legend.position = "left") +
    scale_color_viridis(name = closeness_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE
                                  #breaks = seq(0, 100 , by = 0.02
                                  ))+ xlab(""),
    
    plot_regr_centrality_gal,
  
    plot_network_closeness_can + theme(legend.position = "left") +
    scale_color_viridis(
    name = closeness_label, 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20)))+ xlab(""),
    plot_regr_centrality_can,
  
  plot_network_closeness_az + theme(legend.position = "left") +
    scale_color_viridis(name =closeness_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 1, barheight = 8,
                                  title.position = "top", label.position = "left",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))),
  plot_regr_centrality_az,
  
  ncol = 2,
  nrow = 3,
  
  labels = LETTERS[1:6],
  
  widths = c(2, 1)
)

plot_network_regr_closeness

ggsave(here::here("c_output/figures/plot_network_regr_closeness.png"), height = 12, width = 12)


```





















```{r}
plot_network_regr_centrality <- ggarrange(
  
      plot_network_degree_gal + theme(legend.position = "top") +
    scale_color_viridis(name = in_degree_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 8, barheight = 1,
                                  title.position = "left", label.position = "bottom",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))),
    
    plot_network_closeness_gal + theme(legend.position = "top") +
    scale_color_viridis(name = closeness_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 8, barheight = 1,
                                  title.position = "left", label.position = "bottom",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))),
  plot_regr_centrality_gal,
  
    plot_network_degree_can + theme(legend.position = "top") +
    scale_color_viridis(
    name = in_degree_label, 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))+
    guides(color = guide_colorbar(barwidth = 8, barheight = 1,
                                  title.position = "left", label.position = "bottom",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))),
  
  plot_network_closeness_can + theme(legend.position = "top") +
    scale_color_viridis(
    name = closeness_label, 
    option = "turbo", 
    labels = function(x) sprintf("%.2f", back_transform_log(x),3))+
    guides(color = guide_colorbar(barwidth = 8, barheight = 1,
                                  title.position = "left", label.position = "bottom",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))) ,
   plot_regr_centrality_can,
  
  plot_network_degree_az + theme(legend.position = "top") +
    scale_color_viridis(name =in_degree_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 8, barheight = 1,
                                  title.position = "left", label.position = "bottom",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))),
  
  plot_network_closeness_az + theme(legend.position = "top") +
    scale_color_viridis(name =closeness_label, option = "turbo")+
    guides(color = guide_colorbar(barwidth = 8, barheight = 1,
                                  title.position = "left", label.position = "bottom",
                                  title.hjust = 0.5,
                                  ticks = FALSE,
                                  breaks = seq(0, 100, by = 20))),
  plot_regr_centrality_az,
  
  ncol = 3,
  nrow = 3,
  
  labels = LETTERS[1:9],
  
  widths = c(1.6, 1.6, 1)
)

plot_network_regr_centrality

ggsave(here::here("c_output/figures/plot_network_regr_centrality.png"), height = 12, width = 14)
```






