---
title: "Floristic Distances of Galapagos' Islands"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    df_print: paged
editor_options:
  chunk_output_type: inline
---


This script reads floristic data from each archipelago, filters and arranges it to save floristic mastrices for:

- All plants
- Thalassochorous plants
- Littoral plants
- Littoral exclusive plants - living in the littoral zone only
- Fleshy fruited plants




```{r}
library('readr')
library('tidyverse')
library('tidyr')
library('permute')
library('lattice')
library('vegan')
```



### Open Data

```{r}
data_Galapagos <- read_csv("../Data/Data_Galapagos/data_galapagos.csv", col_names = FALSE, trim_ws = TRUE)
data_Canaries <- read_csv("../Data/Data_Canaries/data_canaries.csv", col_names = FALSE, trim_ws = TRUE)
data_Azores <- read_csv("../Data/Data_Azores/data_azores.csv", col_names = FALSE)

df_Galapagos <- as.data.frame(data_Galapagos)
df_Canaries <- as.data.frame(data_Canaries)
df_Azores <- as.data.frame(data_Azores)


```



## Smooth the data frame

```{r}
df_Galapagos[is.na(df_Galapagos)] <- 0
df_Galapagos <- unite(df_Galapagos, name_species, c(X2, X3), remove=FALSE)
df_Galapagos <- df_Galapagos %>% select(1,2,5:29)


df_Canaries[is.na(df_Canaries)] <- 0

df_Azores[is.na(df_Azores)] <- 0

```





# Filter floristic data to obtain florisitc matrices

This part requires specific filtering because each floristic dataset is different and we need to take into account specific considerations for certain species in each archipelago


## Galapagos

```{r}
# m = matrix

df_Galapagos$name_species <- gsub('Sesuvium_edmondstonei', 'Sesuvium_edmonstonei', df_Galapagos$name_species)
# Sesuvium edmondstonei has been probably misspelled by the data set. There is a specie called Sesuvium edmonstonei, correctly descrived by Wiggins & Porter (1971), native from Galapagos Islands (Title of vols. 1-6 (1791-1802) Transactions of the Linnean Society; for series 2 see Trans. Linn. Soc. London) (it can also be checked on http://powo.science.kew.org/taxon/235141-2). So we have changed the data set to correct the species spelling.

df_Galapagos$name_species <- gsub('Pycreus_polystachyos', 'Cyperus_polystachyos', df_Galapagos$name_species)
# The specie Pycreus polystachyos, found in our data, is a synonim of Cyperus polystachyos, found in Wiggins & Porter (1971). The taxonomical correct name is Cyperus polystachyos Rottb., Descr. Pl. Rar. 21 (1772), because it is ancient than Pycreus polystachyos P.Beauv., Fl. Oware 2: 48, t. 86 (1816). So we have changed Pycreus polystachyos to Cyperus polystachyos on our data.

df_Galapagos <- subset(df_Galapagos, df_Galapagos$X4 %in% c("1", "habitats<500m")) # We filter only lowland species
df_Galapagos <- subset(df_Galapagos, df_Galapagos$X5 %in% c("Id", "En", "EnQ", "Sub-Origin"))   # From lowland species, we filter only native (and endemic) species




m_all_gal <- df_Galapagos %>% select(2,7:18)

m_thalassochory_gal <- subset(df_Galapagos, df_Galapagos$X25 %in% c("0.5", "1", "THA"))   # From lowland & native species, we filter only thalassochoric species
m_thalassochory_gal <- m_thalassochory_gal %>% select(2,7:18)


m_fleshy_fruits_gal <- subset(df_Galapagos, df_Galapagos$X21 %in% c("Berry", "Berry-like", "Drupe", "Drupaceous", "Drupe-like", "Fruit_type"))   # From lowland species, we filter only species with fleshy fruit
m_fleshy_fruits_gal <- m_fleshy_fruits_gal %>% select(2,7:18)


m_littoral_gal <- subset(df_Galapagos, df_Galapagos$X28 %in% c("presence", "exclusive", "littoral"))    # From lowland species, we filter only species present in littoral
m_littoral_gal <- m_littoral_gal %>% select(2,7:18)

# Cyperus liguhris was probably misspelled by Wiggins & Porter (1971). There is a specie called Cyperus ligularis, native from Galapagos Islands (Jørgensen, P.M. León-Yánez, S. (eds.) (1999) Catalogue of the Vascular Plants of Ecuador. Monographs in Systematic Botany from the Missouri Botanical Garden 75. Missouri Botanical Garden Press, St. Louis, 1181 pp.). So we have changed Cyperus liguhris to Cyperus ligularis.

# Cyperus brevifolius doesn't appear on the data set. This specie is correct and native from Galapagos (Lawesson, J. E., H. Adsersen & P. Bentley. 1987. An updated and annotated check list of the vascular plants of the Galapagos Islands. Rep. Bot. Inst. Univ. Aarhus 16: 1–74.), but it simply doesn't appear on the data.

# Conocarpus erecta was probably misspelled by Wiggins & Porter (1971). There is a specie called Conocarpus erectus native from Galapagos Islands (McMullen, C.K. (1999) Flowering plants of the Galapagos. Cornell University Press, Ithaca and London, 370 pp.) that could be mixed with a synonim of Conocarpus erectus that is:  Terminalia erecta (L.) Baill. (https://www.ipni.org/n/171102-1). So we have changed Conocarpus erecta to Conocarpus erectus.


m_littoral_exclusive_gal <- subset(df_Galapagos, df_Galapagos$X28 %in% c("exclusive", "littoral"))    # From lowland species, we filter only species exclusive from the littoral
m_littoral_exclusive_gal <- m_littoral_exclusive_gal %>% select(2,7:18)

```


#### Filter the negative matrixes

```{r}
# m = matrix

m_all_without_thalassochory_gal <- subset(df_Galapagos, !df_Galapagos$X25 %in% c("0.5", "1"))   # We filter NON thalassochoric species
m_all_without_thalassochory_gal <- m_all_without_thalassochory_gal %>% select(2,7:18)

m_all_without_littoral_gal <- subset(df_Galapagos, !df_Galapagos$X28 %in% c("presence", "exclusive"))   # We filter NON litoral species
m_all_without_littoral_gal <- m_all_without_littoral_gal %>% select(2,7:18)

m_all_without_littoral_exclusive_gal <- subset(df_Galapagos, !df_Galapagos$X28 %in% c("exclusive"))  # We filter NON littoral exclusive species
m_all_without_littoral_exclusive_gal <- m_all_without_littoral_exclusive_gal %>% select(2,7:18)


m_all_without_fleshy_gal <- subset(df_Galapagos, !df_Galapagos$X21 %in% c("Berry", "Berry-like", "Drupe", "Drupaceous", "Drupe-like"))  # We filter NON littoral exclusive species
m_all_without_fleshy_gal <- m_all_without_fleshy_gal %>% select(2,7:18)
```




## Canaries

```{r}
# m = matrix

df_Canaries <- subset(df_Canaries, df_Canaries$X3 %in% c("1", "habitat<500m")) # We filter only lowland species
df_Canaries <- subset(df_Canaries, df_Canaries$X4 %in% c("End", "NS", "NP", "status"))   # From lowland species, we filter only native (and endemic) species




m_all_can <- df_Canaries %>% select(1,6:12)

m_thalassochory_can <- subset(df_Canaries, df_Canaries$X15 %in% c("0.5", "1", "thalassochory"))   # We filter only thalassochoric species
m_thalassochory_can <- m_thalassochory_can %>% select(1,6:12)

m_littoral_can <- subset(df_Canaries, df_Canaries$X5 %in% c("presence", "exclusive", "littoral"))   # We filter only littoral species
m_littoral_can <- m_littoral_can %>% select(1,6:12)

m_littoral_exclusive_can <- subset(df_Canaries, df_Canaries$X5 %in% c("exclusive", "littoral"))   # We filter only littoral species
m_littoral_exclusive_can <- m_littoral_exclusive_can %>% select(1,6:12)
```


### Filter the negative matrixes

```{r}
# m = matrix

m_all_without_thalassochory_can <- subset(df_Canaries, !df_Canaries$X15 %in% c("0.5", "1"))   # We filter NON thalassochoric species
m_all_without_thalassochory_can <- m_all_without_thalassochory_can %>% select(1,6:12)


m_all_without_littoral_can <- subset(df_Canaries, !df_Canaries$X5 %in% c("presence", "exclusive"))   # We filter NON litoral species
m_all_without_littoral_can <- m_all_without_littoral_can %>% select(1,6:12)

m_all_without_littoral_exclusive_can <- subset(df_Canaries, !df_Canaries$X5 %in% "exclusive")  # We filter NON littoral exclusive species
m_all_without_littoral_exclusive_can <- m_all_without_littoral_exclusive_can %>% select(1,6:12)
```


## Azores

```{r}
# m = matrix

m_all_az <- df_Azores %>% select(2,10:18) # m_all are all the natives species that live in lowlands


m_thalassochory_az <- subset(df_Azores, df_Azores$X7 %in% c("0.33", "0.5", "1", "THA"))   # We filter only thalassochoric species
m_thalassochory_az <- m_thalassochory_az %>% select(2,10:18)


m_littoral_az <- subset(df_Azores, df_Azores$X3 %in% c("1", "littoral"))   # We filter only litoral species
m_littoral_az <- m_littoral_az %>% select(2,10:18)

m_littoral_exclusive_az <- subset(df_Azores, df_Azores$X2 %in% c("Crithmum_maritimum", "Juncus_maritimus", "Myosotis_maritima", "Ruppia_maritima", "Sagina_maritima", "Polygonum_maritimum", "name_species"))  # We filter only lttoral species that are nowhere but in the littoral zone.
m_littoral_exclusive_az <- m_littoral_exclusive_az %>% select(2,10:18)
```


### Filter the negative matrixes

```{r}
# m = matrix

m_all_without_thalassochory_az <- subset(df_Azores, !df_Azores$X7 %in% c("0.33", "0.5", "1"))   # We filter NON thalassochoric species
m_all_without_thalassochory_az <- m_all_without_thalassochory_az %>% select(2,10:18)


m_all_without_littoral_az <- subset(df_Azores, !df_Azores$X3 %in% "1")   # We filter NON litoral species
m_all_without_littoral_az <- m_all_without_littoral_az %>% select(2,10:18)

m_all_without_littoral_exclusive_az <- subset(df_Azores, !df_Azores$X2 %in% c("Crithmum_maritimum", "Juncus_maritimus", "Myosotis_maritima", "Ruppia_maritima", "Sagina_maritima", "Polygonum_maritimum"))  # We filter NON littoral exclusive species
m_all_without_littoral_exclusive_az <- m_all_without_littoral_exclusive_az %>% select(2,10:18)
```






# 1. Calculate dissimilarities



Dissimilarity Index (DI) measures the relative separation (high dissimilarity) or closeness (low dissimilarity) between two richness of islands It is a symmetrical measure so that this interpretation can apply to either richness 

- A DI of 0 indicates conditions of total closeness under which richness and their species are distributed in the same proportions across all islands. 
- A DI of 1 indicates conditions of total segregation such that the species are located in completely different islands than other species. 


#### 1.3. Bray-Curtis

It is a simetrical index, it means that takes into account double-zero attributes (co-absences). Bray-Curtis treats in exactly the same way co-absences as any other pair of values
and should be used when the zero state is a valid basis for comparing two objects in the similarity calculation (Greenacre, 2008; Todeschini et al., 2012).

The Formula is 1 - 2* (sum of the lesser values for species both samples share) / total number of species counted at both sites. It is an inverse index due to '1-'. . This measure is very reasonable if we want to take into account ABSENCES of species in our data. However, it does not satisfy the triangle inequality axiom, and hence is not a true distance. It means that the sum of the dissimilarities between islands A and B and islands A and C is not necessarily greater than between islands B and C (Oksanen et al., 2020) (Greenacre, 2008; Post-ResearchGate.

This dissimilarity index ranges from zero, when species are similar among sites, to unity, when no single species is shared among them (Castro-Urgal, R. & Traveset, A., 2014).

In our data, absences are true absences because these species haven't been found in the islands that are marked as absences. Moreover, our matrix have 71% of absences and 29% of presences. So there are probably quite co-absences between islands that we think have to be taken into account to compare islands.




```{r}

# Galapagos


df_all_gal <- Calculate_distmatrix_and_save(
  matrix = m_all_gal,file_path = "Dist_Matices_floras_Galapagos/df_all_gal")

df_thalassochory_gal <- Calculate_distmatrix_and_save(
  matrix = m_thalassochory_gal,file_path = "Dist_Matices_floras_Galapagos/df_thalassochory_gal")

df_fleshy_fruits_gal <- Calculate_distmatrix_and_save(
  matrix = m_fleshy_fruits_gal,file_path = "Dist_Matices_floras_Galapagos/df_fleshy_fruits_gal")

df_littoral_gal <- Calculate_distmatrix_and_save(
  matrix = m_littoral_gal,file_path = "Dist_Matices_floras_Galapagos/df_littoral_gal")

df_littoral_exclusive_gal <- Calculate_distmatrix_and_save(
  matrix = m_littoral_exclusive_gal,file_path = "Dist_Matices_floras_Galapagos/df_littoral_exclusive_gal")

df_all_without_thalassochory_gal <- Calculate_distmatrix_and_save(
  matrix = m_all_without_thalassochory_gal,file_path = 
    "Dist_Matices_floras_Galapagos/df_all_without_thalassochory_gal")

df_all_without_littoral_gal <- Calculate_distmatrix_and_save(
  matrix = m_all_without_littoral_gal,file_path = "Dist_Matices_floras_Galapagos/df_all_without_littoral_gal")

df_all_without_littoral_exclusive_gal <- Calculate_distmatrix_and_save(
  matrix = m_all_without_littoral_exclusive_gal,file_path = 
    "Dist_Matices_floras_Galapagos/df_all_without_littoral_exclusive_gal")

df_all_without_fleshy_gal <- Calculate_distmatrix_and_save(
  matrix = m_all_without_fleshy_gal,file_path = "Dist_Matices_floras_Galapagos/df_all_without_fleshy_gal")




# Canaries


df_all_can <- Calculate_distmatrix_and_save(matrix = m_all_can, file_path = "Dist_Matices_floras_Canaries/df_all_can")

df_thalassochory_can <- Calculate_distmatrix_and_save(
  matrix = m_thalassochory_can,file_path = "Dist_Matices_floras_Canaries/df_thalassochory_can")

df_littoral_can <- Calculate_distmatrix_and_save(
  matrix = m_littoral_can,file_path = "Dist_Matices_floras_Canaries/df_littoral_can")

df_littoral_exclusive_can <- Calculate_distmatrix_and_save(
  matrix = m_littoral_exclusive_can,file_path = "Dist_Matices_floras_Canaries/df_littoral_exclusive_can")

df_all_without_thalassochory_can <- Calculate_distmatrix_and_save(
  matrix = m_all_without_thalassochory_can,file_path =
    "Dist_Matices_floras_Canaries/df_all_without_thalassochory_can")

df_all_without_littoral_can <- Calculate_distmatrix_and_save(
  matrix = m_all_without_littoral_can,file_path =
    "Dist_Matices_floras_Canaries/df_all_without_littoral_can")

df_all_without_littoral_exclusive_can <- Calculate_distmatrix_and_save(
  matrix = m_all_without_littoral_exclusive_can,file_path = 
    "Dist_Matices_floras_Canaries/df_all_without_littoral_exclusive_can")





# Azores


df_all_az <- Calculate_distmatrix_and_save(matrix = m_all_az,file_path = "Dist_Matices_floras_Azores/df_all_az")

df_thalassochory_az <- Calculate_distmatrix_and_save(
  matrix = m_thalassochory_az,file_path = "Dist_Matices_floras_Azores/df_thalassochory_az")

df_littoral_az <- Calculate_distmatrix_and_save(
  matrix = m_littoral_az,file_path = "Dist_Matices_floras_Azores/df_littoral_az")

df_littoral_exclusive_az <- Calculate_distmatrix_and_save(
  matrix = m_littoral_exclusive_az,file_path = 
    "Dist_Matices_floras_Azores/df_littoral_exclusive_az")

df_all_without_thalassochory_az <- Calculate_distmatrix_and_save(
  matrix = m_all_without_thalassochory_az,file_path = 
    "Dist_Matices_floras_Azores/df_all_without_thalassochory_az")

df_all_without_littoral_az <- Calculate_distmatrix_and_save(
  matrix = m_all_without_littoral_az,file_path = "Dist_Matices_floras_Azores/df_all_without_littoral_az")

m_all_without_littoral_exclusive_az <- Calculate_distmatrix_and_save(
  matrix = m_all_without_littoral_exclusive_az,file_path = 
    "Dist_Matices_floras_Azores/df_all_without_littoral_exclusive_az")


```





