---
title: "Analyse correlations between island features and plant percentages"
output: html_notebook
---

This script analyses the correlation between island degree (currents), area, and age and percentages of thalassochorous or littoral species.


```{r}
library(EpiModel)
library(shp2graph)
library(tnet)
library(dplyr)
library(rcompanion)
library(igraph)
library(spaa)
library(textshape)
library(ggpubr)
```



## Obtain N plants and percentage plants per island




#### Total number of plants per archipelago

```{r}
df_all_gal<-header.true(m_all_gal)
df_all_can<-header.true(m_all_can)
df_all_az<-header.true(m_all_az)

n_plants_gal<-length(df_all_gal$Genus_Species)
n_plants_can<-length(df_all_can$name_species)
n_plants_az<-length(df_all_az$name_species)
```


#### Number and percentage of thalassochorous and littoral plants per island

```{r}

Nplants_tha_gal<-Calculate_Nplants_perIsland(df_flora=m_thalassochory_gal)
Nplants_litt_gal<-Calculate_Nplants_perIsland(df_flora=m_littoral_gal)

Nplants_tha_can<-Calculate_Nplants_perIsland(df_flora=m_thalassochory_can)
Nplants_litt_can<-Calculate_Nplants_perIsland(df_flora=m_littoral_can)


Nplants_tha_az<-Calculate_Nplants_perIsland(df_flora=m_thalassochory_az)
Nplants_litt_az<-Calculate_Nplants_perIsland(df_flora=m_littoral_az)



# Galapagos

fun_percentage <- function(i) {
    (i)/n_plants_gal*100
}                                             
                                             
Percentage_plants_tha_gal<-Calculate_Percentage_plants_perIsland(
  df_flora=m_thalassochory_gal)


Percentage_plants_litt_gal<-Calculate_Percentage_plants_perIsland(
  df_flora=m_littoral_gal)



# Canaries


fun_percentage <- function(i) {
    (i)/n_plants_can*100
}  

Percentage_plants_tha_can<-Calculate_Percentage_plants_perIsland(
  df_flora=m_thalassochory_can)

Percentage_plants_litt_can<-Calculate_Percentage_plants_perIsland(
  df_flora=m_littoral_can)



# Azores

fun_percentage <- function(i) {
    (i)/n_plants_az*100
} 


Percentage_plants_tha_az<-Calculate_Percentage_plants_perIsland(
  df_flora=m_thalassochory_az)

Percentage_plants_litt_az<-Calculate_Percentage_plants_perIsland(
  df_flora=m_littoral_az)

```




## Obtain degree (connectivity currents) per island



```{r}
Degree_curr_gal<-calculate_degrees_perIsland(dist_matrix =dist_curr_gal)
Degree_curr_can<-calculate_degrees_perIsland(dist_matrix =dist_curr_can)
Degree_curr_az<-calculate_degrees_perIsland(dist_matrix =dist_curr_az)

```

```{r}
Degree_geo_gal<-calculate_degrees_perIsland(dist_matrix =dist_geo_gal)
Degree_geo_can<-calculate_degrees_perIsland(dist_matrix =dist_geo_can)
Degree_geo_az<-calculate_degrees_perIsland(dist_matrix =dist_geo_az)
```



```{r}
m_thalassochory_gal1<-header.true(m_thalassochory_gal)
m_thalassochory_gal2<-m_thalassochory_gal1[,-1]

m_thalassochory_can1<-header.true(m_thalassochory_can)
m_thalassochory_can2<-m_thalassochory_can1[,-1]

m_thalassochory_az1<-header.true(m_thalassochory_az)
m_thalassochory_az2<-m_thalassochory_az1[,-1]



island_names_gal<-sort(colnames(m_thalassochory_gal2))
Archipelago_name_gal<-rep(c("Galapagos"),times=length(island_names_gal))


island_names_can<-sort(colnames(m_thalassochory_can2))
Archipelago_name_can<-rep(c("Canaries"),times=length(island_names_can))

island_names_az<-sort(colnames(m_thalassochory_az2))
Archipelago_name_az<-rep(c("Azores"),times=length(island_names_az))

```


## Load area data per island

```{r}
df_area_gal<-read.csv("../../Data/Data_Galapagos/area_islands_galap.csv")
df_area_can<-read.csv("../../Data/Data_Canaries/area_islands_canaries.csv")
df_area_az<-read.csv("../../Data/Data_Azores/area_islands_azores.csv")

df_area_can1<-sort(df_area_can)
df_area_az1<-sort(df_area_az)

area_vector_gal<-df_area_gal$area
area_vector_can<-df_area_can1$area
area_vector_az<-df_area_az1$area
```



## Load age data per island


```{r}
df_age_gal<-read.csv("../../Data/Data_Galapagos/df_ages_Galap.csv",sep=";")
df_age_can<-read.csv("../../Data/Data_Canaries/df_ages_canaries.csv", sep=";")
df_age_az<-read.csv("../../Data/Data_Azores/df_ages_azores.csv", sep=";")

df_age_can1<-sort(df_age_can)
df_age_az1<-sort(df_age_az)

age_vector_gal<-df_age_gal$age
age_vector_can<-df_age_can1$age
age_vector_az<-df_age_az1$age
```



## Build dataframe with current degree, area, age, number and percentage of plants per island


```{r}


df_perc_N_deg_gal<-Create_dataframe_IslName_perc_N_degree(
  Archipelago_name_vector = Archipelago_name_gal,
  island_names_vector = island_names_gal,
  Percentage_plants_vector1 = Percentage_plants_tha_gal,
  Percentage_plants_vector2 = Percentage_plants_litt_gal,
  Nplants_vector1 = Nplants_tha_gal,
  Nplants_vector2 = Nplants_litt_gal,
  Degree_vector1 = Degree_curr_gal, 
  Degree_vector2 = Degree_geo_gal, 
  area_vector = area_vector_gal,
  age_vector = age_vector_gal)

df_perc_N_deg_can<-Create_dataframe_IslName_perc_N_degree(
  Archipelago_name_vector = Archipelago_name_can,
  island_names_vector = island_names_can,
  Percentage_plants_vector1 = Percentage_plants_tha_can,
  Percentage_plants_vector2 = Percentage_plants_litt_can,
  Nplants_vector1 = Nplants_tha_can,
  Nplants_vector2 = Nplants_litt_can,
  Degree_vector1 = Degree_curr_can, 
  Degree_vector2 = Degree_geo_can, 
  area_vector = area_vector_can,
  age_vector = age_vector_can)

df_perc_N_deg_az<-Create_dataframe_IslName_perc_N_degree(
  Archipelago_name_vector = Archipelago_name_az,
  island_names_vector = island_names_az,
  Percentage_plants_vector1 = Percentage_plants_tha_az,
  Percentage_plants_vector2 = Percentage_plants_litt_az,
  Nplants_vector1 = Nplants_tha_az,
  Nplants_vector2 = Nplants_litt_az,
  Degree_vector1 = Degree_curr_az, 
  Degree_vector2 = Degree_geo_az, 
  area_vector = area_vector_az,
  age_vector = age_vector_az)

df_archipelagos_perc_N_deg<-rbind(df_perc_N_deg_gal,
                                  df_perc_N_deg_can,
                                  df_perc_N_deg_az)


```




*Connection between Fuerteventura y Lanzarote has an error value in the distance matrix, so gives an unrealistic degree. This needs to be corrected*


```{r}

trial<-df_trial<-df_archipelagos_perc_N_deg
df_trial<-df_archipelagos_perc_N_deg[-c(14,18),] # eliminate error data of Lanzarote and Fuerteventura


scatter_curr_degree<-ggscatter(df_trial, x = "Degree_curr", y = "Percentage_plants_tha",color="Archipelago", size=3.5,
                               add = "reg.line", conf.int = TRUE, 
                               cor.coef = TRUE, cor.method = "pearson",
                               xlab = "Degree_curr", ylab = "Percentage_plants_tha",
                               add.params = list(color = "red", fill = "lightgray"))

scatter_geo_degree<-ggscatter(trial, x = "Degree_geo", y = "Percentage_plants_tha",
                              color="Archipelago", size=3.5,
                              add = "reg.line", conf.int = TRUE, 
                              cor.coef = TRUE, cor.method = "pearson",
                              xlab = "Degree_geo", ylab = "Percentage_plants_tha",
                              add.params = list(color = "red", fill = "lightgray"))

scatter_area_degree<-ggscatter(trial, x = "area", y = "Percentage_plants_tha",
                               color="Archipelago", size=3.5,
                               add = "reg.line", conf.int = TRUE, 
                               cor.coef = TRUE, cor.method = "pearson",
                               xlab = "area", ylab = "Percentage_plants_tha",
                               add.params = list(color = "red", fill = "lightgray"))

scatter_age_degree<-ggscatter(trial, x = "age", y = "Percentage_plants_tha", 
                              color="Archipelago", size=3.5,
                              add = "reg.line", conf.int = TRUE, 
                              cor.coef = TRUE, cor.method = "pearson",
                              xlab = "age", ylab = "Percentage_plants_tha",
                              add.params = list(color = "red", fill = "lightgray"))


tha_scatter_archipelagos<-ggarrange(scatter_curr_degree, scatter_geo_degree, scatter_area_degree, scatter_age_degree + rremove("x.text"),
                                    ncol = 2, nrow = 2)
annotate_figure(tha_scatter_archipelagos, 
                top = text_grob("thalassochorous",
                                color = "red", face = "bold", size = 14))
```

```{r}

trial_litt<-df_trial<-df_archipelagos_perc_N_deg
df_trial_litt<-df_archipelagos_perc_N_deg[-c(14,18),]

scatter_litt_curr_degree<-ggscatter(df_trial_litt, x = "Degree_curr", y = "Percentage_plants_litt",
                                    color="Archipelago", size=3.5,
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Degree_curr", ylab = "Percentage_plants_litt",
          add.params = list(color = "red", fill = "lightgray"))

scatter_litt_geo_degree<-ggscatter(trial_litt, x = "Degree_geo", y = "Percentage_plants_litt", 
                                   color="Archipelago", size=3.5,
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Degree_geo", ylab = "Percentage_plants_litt",
          add.params = list(color = "red", fill = "lightgray"))

scatter_litt_area_degree<-ggscatter(trial_litt, x = "area", y = "Percentage_plants_litt", 
                                    color="Archipelago", size=3.5,
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "area", ylab = "Percentage_plants_litt",
          add.params = list(color = "red", fill = "lightgray"))

scatter_litt_age_degree<-ggscatter(trial_litt, x = "age", y = "Percentage_plants_litt",
                                   color="Archipelago", size=3.5,
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "age", ylab = "Percentage_plants_litt",
          add.params = list(color = "red", fill = "lightgray"))



litt_scatter_archipelagos<-ggarrange(scatter_litt_curr_degree, scatter_litt_geo_degree, scatter_litt_area_degree, scatter_litt_age_degree + rremove("x.text"),
          ncol = 2, nrow = 2)
annotate_figure(litt_scatter_archipelagos, top = text_grob("littoral", 
               color = "red", face = "bold", size = 14))
```











